---
title: "Cox assumptions"
format: html
editor: visual
---

# Proportional hazards assumption

```{r,warning=FALSE, message=FALSE}
#| echo: false
#| output: false

library("survminer")
library("survival")
library("broom")
library("tidyverse")

sas <-
  read.csv(
    here::here("data/ss_earlylife_popu.csv"),
    header = TRUE,
    stringsAsFactors = TRUE,
    fileEncoding = "latin1"
  )

sas$ident_EOF=ifelse(sas$ident_EOF==sas$ident_ageexact_bl,sas$ident_ageexact_bl+0.1,sas$ident_EOF)
all_var_list=c("ses_race",
       "an_bmi_cat",
       "ph_MET_wk_current_cat",
       "alc_cat",
       "smok_status",
       "ses_edu_level",
       "ses_income",
       "illness_thyroid_all",
       "fam_history_DTC",
       
       "ui_diab_all_mother",
       #ui_diab_prior_mother,
       #ui_diab_preg_mother,
       
       "ui_hypertension_all",
       #ui_ges_hypertension,
       #ui_eclampsia,
       "ui_epilepsy_prior",
       "ui_vomit",
       "ui_DES",
       "ui_farm_all",
       "ui_mother_smok",
       "ui_hh_smok",
       "ui_mother_age_cat",
       
       "ui_diab_father",
       "ui_father_smok",
       "ui_father_age_cat",
       
       "ui_maternal_1st_born",
       "ui_preg_order_cat",
       "ui_preg_last_cat",
       "ui_multiple_birth",
       "ui_birth_weight_cat",
       "ui_early_gest",
       "ui_breastfed",
       "ui_soy")

covariates=c("ses_race",
       "an_bmi_cat",
       "ph_MET_wk_current_cat",
       "alc_cat",
       "smok_status",
       "ses_edu_level",
       "ses_income",
       "illness_thyroid_all",
       "fam_history_DTC")

surv="Surv(ident_ageexact_bl, ident_EOF, ident_DTC)"
####Model that includes all covariates
all=reformulate(all_var_list, response = surv)%>%coxph(method = "breslow", data=sas)


```

## **Testing proportional Hazards assumption: scaled Schoenfeld residuals**

```{r,warning=FALSE, message=FALSE}
#| echo: false
#| output: true
#| warning: false

###plots of scaled Schoenfeld residuals
for (i in covariates){
  p=reformulate(i, response = surv)%>%coxph(method = "breslow", data=sas)%>%cox.zph()%>%ggcoxzph()
  q1=p$`1`%>%ggplot_build()
  q1$plot$labels$y = paste0("Beta(t) for ",i,sep="")
  q1%>%ggplot_gtable()%>%plot()
}

```

## **Testing proportional Hazards assumption: p-value**

```{r,warning=FALSE, message=FALSE}
#| echo: false
#| output: true
#| warning: false

###Testing proportional Hazards assumption: p-value
all%>%cox.zph()

```

## **Model selection based on AIC and BIC**

```{r,warning=FALSE, message=FALSE}
#| echo: false
#| output: true
#| warning: false
model_select_preliminary = as.data.frame(NULL)
for (i in covariates) {
  model_select = rbind(
    model_select_preliminary,
    reformulate(i, response = surv) %>% coxph(method = "breslow", data = sas) %>%
      glance() %>% select(AIC, BIC) %>% add_column(variable = i, .before = "AIC")
  ) %>% arrange(AIC)
}

print(model_select_preliminary)


potential_var=c(NULL,"ph_MET_wk_current_cat",
       "alc_cat",
       "smok_status",
       "ses_edu_level",
       "ses_income",
       
       "fam_history_DTC")


model_select=reformulate(cbind("illness_thyroid_all", "an_bmi_cat", "ses_race"), response = surv) %>% coxph(method = "breslow", data = sas) %>%
      glance() %>% select(AIC, BIC) %>% add_column(variable = "base model", .before = "AIC")
for (i in potential_var) {
  model_select = rbind(
    model_select,
    reformulate(cbind(i,"illness_thyroid_all", "an_bmi_cat", "ses_race"), response = surv) %>% coxph(method = "breslow", data = sas) %>%
      glance() %>% select(AIC, BIC) %>% add_column(variable = i, .before = "AIC")
  ) %>% arrange(AIC)
}
print(model_select)

```

# Final decision

13/09/2023: Decide to include smoking status anyway since that is a kinda established risk factors

**14/09/2023: After discussing with Cari, decide to include income and levels of education, too, despite the violation of non-proportional hazards assumption as they are important factors to account for detection bias and they are covariates, not the variables of interest in this study.**

**Final list of covariates: "illness_thyroid_all", "an_bmi_cat", "ses_race", "smok_status", "ses_edu_level", "ses_income"**
