---
title: In-utero and infancy factors and thyroid cancer risk in adult women in the Sister Study cohort
toc: true
toc-expand: true
number-sections: true
anchor-sections: true
format:
  html:
    grid:
      sidebar-width: 50px
      margin-width:  200px
      body-width:	   2500px
    embed-resources: true
    code-fold: true
    code-summary: "Show the code"
    code-tools: true
    number_sections: true
date: 'last-modified'
date-format: '[This version:] MMMM DD, YYYY'
execute: 
  warning: false
  output: true
  message: false
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r}
#Time the code running process
start_time <- Sys.time()
```

```{r}
#| label: function
source(paste(here::here(),'/analysis/R_created_functions.R',sep = ""),local = knitr::knit_global())
```

```{r}
#| label: setup-library-and-data
source(paste(here::here(),'/analysis/setup_library_and_data.R',sep = ""),local = knitr::knit_global())
```

```{r}
#| label: var-list-utero
###Create lists of exposure and covariates to avoid repetition in the code

#Covariates to adjust in Cox models
covariates_label_utero = list(
  "ses_race"~"Race"
  # "an_bmi_cat"~"BMI",
  # "smok_status"~"Smoking status",
  # "illness_thyroid_all"~"Benign thyroid disease",
  # "ses_edu_level"~"Personal educational level",
  # "ses_income"~"Household annual income",
  # "ses_ADI_baseline_cat"~"Area deprivation index"
)
covariates_utero = substr(covariates_label_utero, 2, unlist(gregexpr("~", covariates_label_utero)) - 3)

#In-utero and infancy factors
var_label_utero = list(
  "ui_diab_all_mother"~"Pre-pregnancy or gestational diabetes",
  "ui_hypertension_all"~"Gestational hypertension or hypertension-related disorders",
###"ui_epilepsy_prior"~"Epilepsy",                                            #not included because of number of cases<5
  "ui_vomit"~"Morning sickness with vomiting",
###"ui_DES"~"DES use",                                                        #not included because of number of cases<5
  "ui_mother_smok"~"Smoking during pregnancy",
  "ui_hh_smok"~"Household secondhand smoking during pregnancy",
  "ui_preg_last_cat"~"Time since last pregnancy",
  #"ui_preg_last_cont"~"Time since last pregnancy (per year, continuous)",
  "ui_mother_age_cat"~"Age at delivery",
  "ui_mother_age_cont"~"Age at delivery (year, continuous)",
  #"ui_diab_father"~"Diabetes",
  "ui_father_smok"~"Smoking during 3 months before conception",
  "ui_father_age_cat"~"Age at delivery",
  "ui_father_age_cont"~"Age at delivery (year, continuous)",
  "ui_birth_weight_cat"~"Birth weight",
  "ui_birth_weight_cont"~"Birth weight (continuous)",
  #"ui_maternal_1st_born"~"First born", # Accounted for Cari's comment on 02/11/24: I would delete this (first born, yes or no) since it is not showing anything different from the expanded variable. 
  "ui_preg_order_cat"~"Birth order",
  "ui_preg_order_cont"~"Birth order (per birth, continuous)",
  "ui_multiple_birth"~"Multiple birth",
  "ui_early_gest"~"Gestational length at birth",
###"ui_early_gest_detail"~"Gestational age at birth",                         #for exploratory purposes only, number of cases<5
  "ui_breastfed"~"Ever breastfed",
  "ui_breastfed_time"~"Duration of breastfeeding"
###"ui_soy"~"Ever fed soy formula"                                            #not included because of number of cases<5)
)
var_name_utero = substr(var_label_utero, 2, unlist(gregexpr("~", var_label_utero)) - 3)

```

# Main Tables and Figures

## Figure 1: **Flowchart of study population**

```{r}
#| label: flowchart
#| out-width: "50%"

eligible <- boxGrob(glue("Female participants", "in the Sister Study cohort",
                         "n = 50884",
                         pop = txtInt(1032),
                         .sep = "\n"),y = unit(0.5, "npc"),
  x = unit(0.2, "npc"))
included <- boxGrob(glue("Study population",
                         "n = 47913",
                         incl = txtInt(122),
                         .sep = "\n"),y = unit(0.5, "npc"),
  x = unit(0.2, "npc"))

excluded <- boxGrob(glue("Excluded:",
                         " - Any history of invasive cancer prior to baseline: n = 2911",
                         " - Self-reported chemotherapy or radiotherapy for cancer prior",
                         " to baseline: n = 55",
                         " - Withdrawal from the Sister Study: n = 5",
                         tot = 30,
                         uninterested = 12,
                         contra = 30 - 12,
                         .sep = "\n"),
                    just = "left")

grid.newpage()
vert <- spreadVertical(eligible = eligible,
                       included = included)

excluded <- moveBox(excluded,
                    x = 0.65,
                    y =coords(vert$included)$top + distance(vert$eligible, vert$included, half = FALSE, center = FALSE))

for (i in 1:(length(vert) - 1)) {
  connectGrob(vert[[i]], vert[[i + 1]], type = "vert") %>%
    print
}

connectGrob(vert$eligible, excluded, type = "L")

# Print boxes
print(vert)
print(excluded)
rm(eligible,excluded,included,vert)

```

## Table 1: **Characteristics of the study population (n=47810)**

```{r}
#| label: tbl-descript
#| tbl-cap: "Characteristics of the study population (n=47810)"
#| cache: true
#| dependson: c("setup-library-and-data", "var-list-utero", "functions")
#| cache.extra: !expr file.mtime("C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/data/ss_earlylife_popu.csv")

table1_descript_var_label = list(
  "ident_ageexact_bl"~"Age at baseline",
  "ses_race"~"Race",
  "an_bmi_cat"~"BMI",
#  "smok_status"~"Smoking status",
  "illness_thyroid_all"~"Benign thyroid disease",
  "ses_edu_level"~"Educational level",
  "ses_income"~"Household annual income"
#  "ses_ADI_baseline_cat"~"Area deprivation index"
)
table1_descript_var_name = substr(table1_descript_var_label, 2, unlist(gregexpr("~", table1_descript_var_label)) - 3)

descript_covar = descript_table(var_name = table1_descript_var_name,
                                var_label = table1_descript_var_label,
                                data = earlylife_popu) |> sort_birth_weight()
rm(table1_descript_var_label,table1_descript_var_name)
descript_covar
```

## Table 2: **Association between in-utero and newborn factors and thyroid cancer incidence in the Sister Study cohort**

```{r}
#| label: utero
#| output: false
#| cache: true
#| dependson: c("setup-library-and-data", "var-list-utero", "functions")
#| cache.extra: !expr file.mtime("C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/data/ss_earlylife_popu.csv")

###To create Table Association between in-utero and infancy factors and thyroid cancer incidence

####Loop over in utero and infancy factors in cox models, both univariable and multivariable analyses
descript_utero = descript_table(var_name = var_name_utero, 
                                var_label = var_label_utero,
                                data = earlylife_popu)|> sort_birth_weight()

# univariable_utero=cox_model(surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
#                      var_name = var_name_utero,
#                      var_label = var_label_utero,
#                      covar = NULL,
#                      data=earlylife_popu)|> sort_birth_weight()

multivariable_utero=cox_model(surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
                       var_name = var_name_utero,
                       var_label = var_label_utero,
                       covar = covariates_utero,
                       data=earlylife_popu)|> sort_birth_weight()

#assemble descriptive table and cox models' results
cox_utero = tbl_merge(
  tbls = list(descript_utero, 
              #univariable_utero, 
              multivariable_utero),
  tab_spanner = c("**N (%)**",
                  #"**Univariable models**",
                  "**Multivariable models**")) |> 
  #modify_column_hide(c(n_event_3,exposure_3))|>
  polish_utero(group_name = TRUE)


```

```{r}
#| label: tbl-utero
#| tbl-cap: "Association between in-utero and infancy factors and thyroid cancer incidence"
#| cache: true
#| dependson: c("setup-library-and-data", "var-list-utero", "functions","tbl-utero")
#| cache.extra: !expr file.mtime("C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/data/ss_earlylife_popu.csv")
#return the table 
cox_utero
#Exclude first born 
descript_utero_preg_last_cont = descript_table(var_name = "ui_preg_last_cont",
                       var_label = list("ui_preg_last_cont"~"Time since last pregnancy (per year, continuous)"),
                                data=earlylife_popu[earlylife_popu$ui_preg_last_cont>0,])
multivariable_utero_preg_last_cont=cox_model(surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
                       var_name = "ui_preg_last_cont",
                       var_label = list("ui_preg_last_cont"~"Time since last pregnancy (per year, continuous)"),
                       covar = covariates_utero,
                   data=earlylife_popu[earlylife_popu$ui_preg_last_cont>0,])|> sort_birth_weight()
descript_utero_preg_last_cont
multivariable_utero_preg_last_cont
```

## Table 3: **Simultaneously adjusted association between thyroid cancer risk and pre-pregnancy or gestational diabetes, gestational hypertension or hypertension-related disorders, birth weight (categorical variable), and gestational age at birth**

```{r}
#| label: tbl-hyper-diab-birthweight-cat
#| tbl-cap: "Simultaneously adjusted association between thyroid cancer risk and pre-pregnancy or gestational diabetes, gestational hypertension or hypertension-related disorders, birth weight, and gestational age at birth"

surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)"
mutual_label_cat = list(
  "ui_diab_all_mother" ~ "Pre-pregnancy or gestational diabetes",
  "ui_hypertension_all" ~ "Gestational hypertension or pre-eclampsia",
  "ui_early_gest" ~ "Gestational age at birth",
  "ui_birth_weight_cat" ~ "Birth weight"
)
mutual_name_cat = substr(mutual_label_cat, 2, unlist(gregexpr("~", mutual_label_cat)) - 3)

hyper_diab_birthweight_mutual_cat = reformulate(c(mutual_name_cat, covariates_utero), response = surv) |>
  coxph(method = "breslow", data = earlylife_popu) |>
  tbl_regression(
    exponentiate = TRUE,
    conf.int = TRUE,
    conf.level = 0.95,
    add_estimate_to_reference_rows = TRUE,
    label = mutual_label_cat,
    include = mutual_name_cat
  )|> sort_birth_weight()|>
  modify_header(n_event ~ "**Number of events**",
                exposure ~ "**Person-years**",
                estimate ~ "**HR**") |>
  modify_table_body( ~ .x |> dplyr::relocate(exposure, .before = estimate))|>
  modify_footnote(abbreviation = TRUE,
                         `estimate` ~ "Multivariable models were adjusted for attained age (timescale), and self-identified race/ethnicity")

hyper_diab_birthweight_mutual_cat

```

# Supplementary materials

## Supplementary table 1: **Association between birth weight and thyroid cancer risk, after excluding participants who were born at least 2 weeks before the due date or those who were part of a multiple birth**

```{r}
#| label: sensi-birthweight
#| output: false
surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)"

non_preterm = earlylife_popu |> filter(ui_early_gest != "Born at least 2 weeks before the due date" & (ui_multiple_birth != "Multiple birth" ))


birth_weight_name = c("ui_birth_weight_cat", "ui_birth_weight_cont")
birth_weight_label = list(
  "ui_birth_weight_cat" ~ "Birth weight",
  "ui_birth_weight_cont" ~ "Birth weight (continuous)"
)

descript_birth = descript_table(var_name = birth_weight_name,
                                var_label = birth_weight_label,
                                data = non_preterm)

multivariable_birth = cox_model(
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
  var_name = birth_weight_name,
  var_label = birth_weight_label,
  covar = covariates_utero,
  data = non_preterm
)

#assemble descriptive table and cox models' results
cox_birth_weight = tbl_merge(
  tbls = list(descript_birth, multivariable_birth),
  tab_spanner = c("**N (%)**",
                  "**Multivariable models**")
) |> sort_birth_weight() |> polish_utero()
```

```{r}
#| label: tbl-sensi-birthweight
#| tbl-cap: "Association between birth weight and thyroid cancer risk, after excluding participants who were born more than 2 weeks before due date or those who were part of a multiple birth"
cox_birth_weight
```

## Supplementary table 2: **Association between birth weight (continuous) and thyroid cancer incidence in the Sister Study cohort by baseline examiner-measured BMI, Area Deprivation Index, personal educational level, and household annual income**

```{r}
#| label: tbl-birthweight-strat
#| tbl-cap: "Association between birth weight (continuous) and thyroid cancer incidence in the Sister Study participants by baseline examiner-measured BMI, Area Deprivation Index, levels of education, and household annual income"

strat_BMI_birthweight = stratification_long(
  strat_var = an_bmi_cat,
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
  var_name = "ui_birth_weight_cont",
  var_label = list("ui_birth_weight_cont" ~ "Birth weight (continuous)"),
  covar = covariates_utero,
  data = earlylife_popu
)

strat_benign_thyroid_birthweight = stratification_long(
  strat_var = illness_thyroid_all,
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
  var_name = "ui_birth_weight_cont",
  var_label = list("ui_birth_weight_cont" ~ "Birth weight (continuous)"),
  covar = covariates_utero,
  data = earlylife_popu
)


strat_edu_birthweight = stratification_long(
  strat_var = ses_edu_level,
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
  var_name = "ui_birth_weight_cont",
  var_label = list("ui_birth_weight_cont" ~ "Birth weight (continuous)"),
  covar = covariates_utero,
  data = earlylife_popu
)

strat_income_birthweight = stratification_long(
  strat_var = ses_income,
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
  var_name = "ui_birth_weight_cont",
  var_label = list("ui_birth_weight_cont" ~ "Birth weight (continuous)"),
  covar = covariates_utero,
  data = earlylife_popu
)


strat_all_birthweight = tbl_stack(list(
  strat_BMI_birthweight,
  strat_benign_thyroid_birthweight,
  #strat_ADI_birthweight,
  strat_edu_birthweight,
  strat_income_birthweight
))

strat_all_birthweight = strat_all_birthweight |> modify_table_body(
  ~ .x  |>
    dplyr::add_row(
      label = "Examiner-measured BMI",
      .before = grep("<25", strat_all_birthweight$table_body$label)
    )  |>
    dplyr::add_row(
      label = "Benign thyroid disease",
      .before = grep(
        "No", 
        strat_all_birthweight$table_body$label
        ) + 1
    )  |>
    dplyr::add_row(
      label = "Personal educational level",
      .before = grep(
        "Highschool or GED or less",
        strat_all_birthweight$table_body$label
      ) + 2
    )  |>
    dplyr::add_row(
      label = "Household annual income",
      .after = grep(
        "Bachelor's degree or higher",
        strat_all_birthweight$table_body$label
      ) + 3
    )
) |>
  modify_table_styling(
    column = "label",
    rows = strat_all_birthweight$table_body$label %in% c(
      "Examiner-measured BMI",
      "Benign thyroid disease",
      "Personal educational level",
      "Household annual income"
    ),
    text_format = "bold"
  )   

strat_all_birthweight
p_interaction_BMI_birthweight=p_interaction(
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
  var_name1 = "ui_birth_weight_cont",
  var_name2 = "an_bmi_cat",
  covar = covariates_utero,
  data = earlylife_popu
)
p_interaction_benign_thyroid_birthweight=p_interaction(
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
  var_name1 = "ui_birth_weight_cont",
  var_name2 = "illness_thyroid_all",
  covar = covariates_utero,
  data = earlylife_popu
)
p_interaction_edu_birthweight=p_interaction(
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
  var_name1 = "ui_birth_weight_cont",
  var_name2 = "ses_edu_level",
  covar = covariates_utero,
  data = earlylife_popu
)

p_interaction_income_birthweight=p_interaction(
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
  var_name1 = "ui_birth_weight_cont",
  var_name2 = "ses_income",
  covar = covariates_utero,
  data = earlylife_popu
)
p_interaction_BMI_birthweight$`Pr(>|Chi|)`[2]
p_interaction_benign_thyroid_birthweight$`Pr(>|Chi|)`[2]
p_interaction_edu_birthweight$`Pr(>|Chi|)`[2]
p_interaction_income_birthweight$`Pr(>|Chi|)`[2]


```

## Supplementary table 3: **Association between in-utero and newborn factors and thyroid cancer incidence in the Sister Study participants, when considering competing risk, only medically confirmed cases, or papillary histology, complete case analysis, and the E-values**

```{r}
#| label: var-list-utero-for-sup-tables
###Create lists of exposure and covariates that have a significant association

#In-utero and infancy factors
var_label_utero_sup = list(
  "ui_diab_all_mother"~"Pre-pregnancy or gestational diabetes",
  "ui_hypertension_all"~"Gestational hypertension or hypertension-related disorders",
  "ui_birth_weight_cat"~"Birth weight",
  "ui_birth_weight_cont"~"Birth weight (continuous)",
  "ui_early_gest"~"Gestational age at birth"
)
var_name_utero_sup = substr(var_label_utero_sup, 2, unlist(gregexpr("~", var_label_utero_sup)) - 3)

```

```{r}
#| label: utero-medically-confirmed
#| output: false
#| cache: true
#| dependson: c("setup-library-and-data", "var-list-utero", "functions")
#| cache.extra: !expr file.mtime("C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/data/ss_earlylife_popu.csv")

cox_medical_confirmed = cox_model(
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC_source)",
  var_name = var_name_utero_sup,
  var_label = var_label_utero_sup,
  covar = covariates_utero,
  data = earlylife_popu,
  model_name = "Medically confirmed cases only"
)|> sort_birth_weight()

```

```{r}
#| label: utero-PTC
#| output: false
#| cache: true
#| dependson: c("setup-library-and-data", "var-list-utero", "functions")
#| cache.extra: !expr file.mtime("C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/data/ss_earlylife_popu.csv")

cox_PTC_utero = cox_model(
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC_PTC)",
  var_name = var_name_utero_sup,
  var_label = var_label_utero_sup,
  covar = covariates_utero,
  data = earlylife_popu,
  model_name = "Papillary thyroid cancer only"
)|> sort_birth_weight()

```

```{r}
#| label: utero-cmprsk
#| output: false
##| cache: true
##| dependson: c("setup-library-and-data", "var-list-utero", "functions")
##| cache.extra: !expr file.mtime("C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/data/ss_earlylife_popu.csv")
 earlylife_popu$ident_compete_factor = factor(earlylife_popu$ident_compete)
 
 cox_cmprsk_utero = competing_risk_model(
   surv = "Surv(ident_ageexact_bl, ident_EOF, ident_compete_factor)",
  var_name = var_name_utero_sup,
  var_label = var_label_utero_sup,
   covar = covariates_utero,
   data = earlylife_popu,
   model_name = "Competing risks"
 )|> sort_birth_weight()

```

```{r}
#| label: utero-complete-case
#| output: false

cox_complete_ges_hypertension = cox_model(
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
  var_name = "ui_hypertension_all",
  var_label = list("ui_hypertension_all"~"Gestational hypertension or hypertension-related disorders"),
  covar = covariates_utero,
  data = earlylife_popu[earlylife_popu$ui_hypertension_all!="Unknown",]
)|> sort_birth_weight()

cox_complete_ges_diab = cox_model(
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
  var_name = "ui_diab_all_mother",
  var_label = list("ui_diab_all_mother"~"Pre-pregnancy or gestational diabetes"),
  covar = covariates_utero,
  data = earlylife_popu[earlylife_popu$ui_diab_all_mother!="Unknown",]
)|> sort_birth_weight()

cox_complete_birth_weight_cat = cox_model(
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
  var_name = "ui_birth_weight_cat",
  var_label = list("ui_birth_weight_cat"~"Birth weight"),
  covar = covariates_utero,
  data = earlylife_popu[earlylife_popu$ui_birth_weight_cat!="Unknown",]
)|> sort_birth_weight()


cox_complete_birth_weight_cont = cox_model(
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
  var_name = "ui_birth_weight_cont",
  var_label = list("ui_birth_weight_cont"~"Birth weight (continuous)"),
  covar = covariates_utero,
  data = earlylife_popu[!is.na(earlylife_popu$ui_birth_weight_cont),]
)|> sort_birth_weight()


cox_complete_early_gest = cox_model(
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
  var_name = "ui_early_gest",
  var_label = list("ui_early_gest"~"Gestational length at birth"),
  covar = covariates_utero,
  data = earlylife_popu[earlylife_popu$ui_early_gest!="Unknown",]
)|> sort_birth_weight()

cox_complete = tbl_stack(list(cox_complete_ges_diab, cox_complete_ges_hypertension, cox_complete_birth_weight_cat, cox_complete_birth_weight_cont, cox_complete_early_gest))|>
  polish_utero()
```

```{r}
#| label: e-values
#| output: false
#| cache: true
#| dependson: c("data", "var-list-utero", "functions")
#| cache.extra: !expr file.mtime("C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/data/ss_earlylife_popu.csv")
multivariable_evalues=multivariable_utero|>modify_table_body(filter,variable %in% c("ui_diab_all_mother",   "ui_hypertension_all","ui_birth_weight_cat","ui_birth_weight_cont","ui_early_gest"))|>polish_utero(group_name = TRUE)
Evalues=cbind(multivariable_evalues$`_data`$estimate,multivariable_evalues$`_data`$conf.low,multivariable_evalues$`_data`$conf.high)

colnames(Evalues)=c("estimate","conf.low","conf.high")

Evalues=round(add_evalues(Evalues),2)
Evalues=cbind(multivariable_evalues$`_data`$label,Evalues)|>as.data.frame()
```

```{r}
#| label: utero-censor-all
#| ouput: false

sensi_censor_utero = tbl_merge(
  tbls = list(
    cox_cmprsk_utero,
    cox_medical_confirmed,
    cox_PTC_utero
  ),
  tab_spanner = FALSE
)  |>
  modify_column_hide(
    c(
      exposure_1,
      exposure_2,
      exposure_3
    )
  )|>  polish_utero(group_name = TRUE)
```

```{r}
#| label: tbl-utero-censor-all
#| tbl-cap: "Association between in-utero and infancy factors and thyroid cancer incidence in the Sister Study participants, when considering only medically confirmed cases, papillary histology, competing risks, or when stratified by tumor size, and cancer stage"
sensi_censor_utero
cox_complete
Evalues
```

```{r}
stop_time=Sys.time() - start_time
print("Time to run the analyses: ")
print(stop_time)
```
