---
title: Childhood and adolescent factors and thyroid cancer risk in adult women in the Sister Study cohort
toc: true
toc-expand: true
number-sections: true
anchor-sections: true
format:
  html:
    embed-resources: true
    code-fold: true
    code-summary: "Show the code"
    code-tools: true
execute: 
  warning: false
  output: true
editor: visual
editor_options: 
  chunk_output_type: console
---

## Methods

### Study population:

The Sister Study is a prospective cohort study aimed at studying environmental and genetic determinants of breast cancer and other chronic diseases. Details of the study design, data collection, and outcome measurements are described elsewhere. Briefly, between 2003 and 2009, the study recruited a total of 50884 women aged 35 to 74 residing in the United States and Puerto Rico who had a sister diagnosed with breast cancer but no prior diagnosis with the disease themselves. At baseline, participants provided detailed information on demographic, medical, lifestyle, and reproductive factors through computer-assisted telephone interviews and written questionnaires, including questions about potential maternal, in utero and infancy exposures. In-person examinations were also conducted to collect anthropometric measurements and biological samples. After the baseline, participants are contacted annually for basic health information updates and every 2-3 years for more detailed follow-up questionnaires, and when necessary, through telephone calls, e-mails, or correspondence to retrieve information on incident cancer and other health outcomes. The response rates to the follow-up questionnaires reached 85% or more (refs). Agreement between self-report and medical record cancer diagnosis was excellent (positive predictive value = xxxxx%) (ref, found in Elisa's paper). Self-reported cancer was confirmed by medical records/pathology report when available. Deaths were ascertained through linkage of the cohort to the National Death Index. The Sister Study was approved by National Institute of Environmental Health Services/NIH and Copernicus Group Institutional Review Boards, and all participants provided written informed consent.

In the current study, we excluded women who withdrew from the study (n=5), who had any history of invasive cancer prior to baseline (n=2911), of total thyroidectomy (n=103), and who reported undergoing chemotherapy or radiotherapy for cancer prior to baseline (n=55). After exclusions, we included 47810 individuals in our study population. The follow-up time was accounted from the date of enrollment to the date of the first thyroid cancer or other invasive cancer diagnosis, death, loss of follow-up, or censoring date, whichever occurred first.

### Thyroid cancer definition:

We considered only well-differentiated thyroid cancer cases and classified them into different histological categories: Papillary thyroid cancer, follicular thyroid cancer and unspecified histology (Supplementary table 1). We censored cases recorded as moderately (n=3) or poorly differentiated (n=3), as well as cases with a histology code of other cancer types, e.g., anaplastic (n=) and medullary cancer (n=) at the date of diagnosis.

### Exposure definition:

### Covariates: 

In-person measurements of weight and height at baseline were used to calculate body mass index (BMI, the weight in kilograms divided by the square of the height in meters). Information on other baseline characteristics, including age, self-identified race/ethnicity, smoking status, the highest level of education, household income, and a personal history of benign thyroid disease were assessed during the telephone interview. Individuals were considered to have a baseline history of benign thyroid disease if they reported hyperthyroidism, hypothyroidism, non-toxic goiter/thyroid nodules, any other thyroid conditions, or if they ever used thyroid hormone substitutes, antithyroid, or iodine/iodide medication. We also considered the Area Deprivation Index (ADI) rankings, which was constructed based on 17 weighted US census income, educational attainment, employment and housing indicators (eg, unemployment rate and median home value) obtained from the American Community Survey. Each neighbourhood received a percentile ranking, with a higher percentage of ADI indicating a higher level of deprivation. The 2000 ADI rankings were linked to each participants' enrolment residential address using Census block-group identification codes. Participants were divided into four groups based on approximate ADI quartiles within the cohort.

```{r}
#| label: setup
###Necessary R library and data loading
library(here)
library(tidyverse)
library(gtsummary)
library(survival)
library(cmprsk)
library(broom)
library(broom.helpers)
library(tidymodels)
library(purrr)

directory="C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS"

earlylife_popu =
  read.csv(paste(directory,"/data/ss_earlylife_popu.csv",sep = ""),
    #because quarto cannot knit the output document if here::here("data/ss_earlylife_popu.csv") is put inside read.csv like this,
    header = TRUE,
    stringsAsFactors = TRUE,
    fileEncoding = "latin1"
  )

earlylife_popu=earlylife_popu|>
  mutate(ident_DTC_PTC=ifelse(ident_DTC_histo=="2) PTC",1,0),
         ident_DTC_stageI=ifelse(ident_DTC_stage_AJCC7=="2) I",1,ident_DTC),
         ident_DTC_adv_stage=ifelse(ident_DTC_stage_AJCC7=="2) I",0,ident_DTC),
         ident_DTC_1cm=ifelse(ident_DTC_size=="2) <1 cm",1,ident_DTC),
         ident_DTC_large_tumor=ifelse(ident_DTC_size=="2) <1 cm",0,ident_DTC))

source(paste(directory,'/Early-life-Thyroid-cancer-SS/analysis/R_created_functions.R',sep = ""),local = knitr::knit_global())


```

```{r, file='C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/Early-life-Thyroid-cancer-SS/analysis/R_created_functions.R'}
```

# **Paper 2: Association between childhood and adolescent factors and thyroid cancer incidence**

```{r}
#| label: var-list-childhood
###Create lists of exposure and covariates to avoid repetition in the code

#Covariates to adjust in Cox models
covariates_label_childhood = list(
  "ses_race" ~ "Race",
  "an_bmi_cat" ~ "BMI",
  "smok_status" ~ "Smoking status",
  "ses_edu_level" ~ "Educational level",
  "ses_income" ~ "Household income",
  "illness_thyroid_all" ~ "Benign thyroid disease"
)
covariates_childhood = substr(covariates_label_childhood, 2, unlist(gregexpr("~", covariates_label_childhood)) - 3)

# add_covariates_label = list(
#   "rep_meno_age_cat"~"Age at menopause"
#   "ph_MET_wk_current_cat"~"Physical activity (MET hours/week)"  ,
#   "ph_MET_wk_current_cont"~"Physical activity (MET hours/week)"  ,
#   "PH_CurrentTotMETHrsPerWeek_CDC"~"Physical activity (MET hours/week)",
#   "ph_hrs_wk_current_cont"~"Physical activity (total hours/week)"    ,
#   "ses_marital_status"~"Marital status"
#   "sm_pack_yr_cat"~"Smoking (pack/year)"            ,
#   "sm_pack_yr_cont"~"Smoking (pack/year)"
#   "sm_yr_smok_cat"~"Years of smoking"    ,
#   "sm_yr_smok_cont"~"Years of smoking"   ,
#   "sm_ETS_ever"~"Ever secondhand smoking"       ,
#   "sm_ETS_total_yr_cat"~"Total years of secondhand smoking"    ,
#   "sm_ETS_total_yr_cont"~"Total years of secondhand smoking"
# )
# add_covariates = substr(add_covariates_label, 2, unlist(gregexpr("~", add_covariates_label)) - 3)

#Childhood factors
var_label_childhood = list(
  "rep_br_dev_age_cat" ~ "Age at breast development"  ,
  "rep_br_dev_age_cont" ~ "Age at breast development" ,
  "rep_age_menarche_cat" ~ "Age at menarche"   ,
  "rep_age_menarche_cont" ~ "Age at menarche"   ,
  "rep_bc_ever" ~ "Ever used hormonal birth control" ,
  "rep_bc_ado" ~ "Hormonal birth control under age 20"  ,
  "rep_bc_ado_dur_cat" ~ "Hormonal birth control duration under age 20",
  "rep_bc_ado_dur_cont" ~ "Hormonal birth control duration under age 20"   ,
  "rep_bc_min_age_cat" ~ "Start age for hormonal birth control" ,
  "rep_bc_min_age_cont" ~ "Start age for hormonal birth control"    ,
  # "rep_proges_ado"~"Progestin-only birth control under age 20"     ,
  # "rep_proges_ado_dur_cat"~"Progestin-only birth control duration under age 20" ,
  # "rep_proges_ado_dur_cont"~"Progestin-only birth control duration under age 20"    ,
  # "rep_proges_min_age_cat"~"Start age for progestin-only birth control"    ,
  # "rep_proges_min_age_cont"~"Start age for progestin-only birth control"   ,
  # "rep_combined_ado"~"Combined hormonal birth control under age 20" ,
  # "rep_combined_ado_dur_cont"~"Combined hormonal birth control duration under age 20" ,
  # "rep_combined_ado_dur_cat"~"Combined hormonal birth control duration under age 20"   ,
  # "rep_combined_min_age_cont"~"Start age for combined hormonal birth control"   ,
  # "rep_combined_min_age_cat"~"Start age for combined hormonal birth control"  ,
  # "rep_bc_ado_type"~"Types of hormonal birth control under age 20"      ,
  # "rep_bc_ingredient"~"Ingredient of hormonal birth control under age 20"   ,
  "an_weight_age10" ~ "Weight at age 10 vs. peers"    ,
  "an_height_age10" ~ "Height at age 10 vs. peers"    ,
  "an_weight_teen" ~ "Teenage weight vs. peers"      ,
  "an_age_tallest" ~ "Age reached full adult height"      ,
  "ph_hrs_wk_child_cat" ~ "Physical activity before age 13 (total hours/week)"   ,
  "ph_hrs_wk_child_cont" ~ "Physical activity before age 13 (total hours/week)"   ,
  "ph_MET_wk_child_cat" ~ "Physical activity before age 13 (MET hours/week)"  ,
  "ph_MET_wk_child_cont" ~ "Physical activity before age 13 (MET hours/week)"   ,
  "ph_hrs_wk_teen_cat" ~ "Physical activity from 13 and 19 (total hours/week)"  ,
  "ph_hrs_wk_teen_cont" ~ "Physical activity from 13 and 19 (total hours/week)"  ,
  "ph_MET_wk_teen_cat" ~ "Physical activity between the age of 13 and 19 (MET hours/week)"    ,
  "ph_MET_wk_teen_cont" ~ "Physical activity between the age of 13 and 19 (MET hours/week)"   ,
  "ph_MET_wk_before20_cat" ~ "Physical activity before age 20 (MET hours/week)"      ,
  "ph_MET_wk_before20_cont" ~ "Physical activity before age 20 (MET hours/week)" ,
  "ph_hrs_wk_before20_cat" ~ "Physical activity before age 20 (total hours/week)"    ,
  "ph_hrs_wk_before20_cont" ~ "Physical activity before age 20 (total hours/week)"    ,
  "alc_drinks_before20_cat" ~ "Alcohol consumption under 20 (drinks/year)"     ,
  "alc_drinks_before20_cont" ~ "Alcohol consumption under 20 (drinks/year)"     ,
  "alc_binge_time_519_cat" ~ "Times per year drank 4+ drinks under 20"     ,
  "alc_binge_time_519_cont" ~ "Times per year drank 4+ drinks under 20"     ,
  "alc_age_start_reg_cat" ~ "Age started drinking regularly"         ,
  "alc_age_start_reg_cont" ~ "Age started drinking regularly"        ,
  "sm_age_start_smok_cat" ~ "Age started smoking"   ,
  "sm_age_start_smok_cont" ~ "Age started smoking"  ,
  "sm_age_stop_smok_cat" ~ "Age stopped smoking"    ,
  "sm_age_stop_smok_cont" ~ "Age stopped smoking"   ,
  "sm_ETS_before18" ~ "Secondhand smoking under 18"     ,
  "sm_ETS_before18_yr_cat" ~ "Total years of secondhand smoking under 18"   ,
  "sm_ETS_before18_yr_cont" ~ "Total years of secondhand smoking under 18"  ,
  "smok_before20" ~ "Smoking before 20"       ,
  "sm_pack_yr_1019_cat" ~ "Smoking from 10 and 19 (Pack-years)"    ,
  "sm_pack_yr_1019_cont" ~ "Smoking from 10 and 19 (Pack-years)"  ,
  "sm_pack_yr_before20_cat" ~ "Smoking before 20 (Pack-years)"  ,
  "sm_pack_yr_before20_cont" ~ "Smoking before 20 (Pack-years)"  ,
  "sm_yr_smok_1019_cat" ~ "Total years of smoking from 10 to 19"    ,
  "sm_yr_smok_1019_cont" ~ "Total years of smoking from 10 to 19"   ,
  "sm_yr_smok_before20_cat" ~ "Total years of smoking before 20"     ,
  "sm_yr_smok_before20_cont" ~ "Total years of smoking before 20"   ,
  "ses_income_child" ~ "Family income while growing up"         ,
  "ses_hunger_child" ~ "Ever not having enough to eat during childhood"          ,
  "ses_edu_level_child" ~ "Highest household education level at age 13",
  "ses_hh_compo_child" ~ "Household composition at age 13",
"ses_urban_child" ~ "Childhood residence"
)
var_name_childhood = substr(var_label_childhood, 2, unlist(gregexpr("~", var_label_childhood)) - 3)

```

##Main tables and figures

### Table 1: Association between childhood and adolescence exposures and thyroid cancer incidence in the Sister Study participants

```{r}
#| label: tbl-stats-childhood
#| tbl-cap: "Association between childhood factors and thyroid cancer incidence"
#| cache: true
#| dependson: c("setup", "var-list-childhood", "functions")
#| cache.extra: !expr file.mtime("C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/data/ss_earlylife_popu.csv")

descript_childhood = descript_table(var_name = var_name_childhood, 
                                var_label = var_label_childhood,
                                data = earlylife_popu)

univariate_childhood=cox_model(surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
                     var_name = var_name_childhood,
                     var_label = var_label_childhood,
                     covar = NULL,
                     data=earlylife_popu)

multivariate_childhood=cox_model(surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
                       var_name = var_name_childhood,
                       var_label = var_label_childhood,
                       covar = covariates_childhood,
                       data=earlylife_popu)

#assemble descriptive table and cox models' results
cox_childhood = tbl_merge(
  tbls = list(descript_childhood, univariate_childhood, multivariate_childhood),
  tab_spanner = c("**N (%)**",
                  "**Univariate models**",
                  "**Multivariate models**")) 

#return the table 
rm(descript_childhood, univariate_childhood, multivariate_childhood)
cox_childhood
```

### Table 2: Association between childhood and adolescent factors and thyroid cancer incidence in the Sister Study participants, by birth year (tertiles), ADI index, educational levels, and household income at baseline

```{r}
#| label: tbl-childhood-strat
#| tbl-cap: "Association between childhood and adolescent factors and thyroid cancer incidence in the Sister Study participants, by birth year (tertiles), ADI index, educational levels, and household income at baseline"
#| cache: true
#| dependson: c("setup", "var-list-childhood", "functions")
#| cache.extra: !expr file.mtime("C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/data/ss_earlylife_popu.csv")

strat_birth_yr_childhood = stratification(
  strat_var = ident_birth_year,
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
  var_name = var_name_childhood,
  var_label = var_label_childhood,
  covar = covariates_childhood,
  data = earlylife_popu
)
strat_ADI_childhood = stratification(
  strat_var = ses_ADI_baseline_cat,
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
  var_name = var_name_childhood,
  var_label = var_label_childhood,
  covar = covariates_childhood,
  data = earlylife_popu
)

strat_edu_childhood = stratification(
  strat_var = ses_edu_level,
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
  var_name = var_name_childhood,
  var_label = var_label_childhood,
  covar = covariates_childhood,
  data = earlylife_popu
)

strat_income_childhood = stratification(
  strat_var = ses_income,
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
  var_name = var_name_childhood,
  var_label = var_label_childhood,
  covar = covariates_childhood,
  data = earlylife_popu
)

tab_spanner = c(
  strat_birth_yr_childhood$table_styling$header$spanning_header |> unique(),
  strat_ADI_childhood$table_styling$header$spanning_header |> unique(),
  strat_edu_childhood$table_styling$header$spanning_header |> unique(),
  strat_income_childhood$table_styling$header$spanning_header |> unique()
) |> na.omit()

strat_all_childhood = tbl_merge(
  tbls = c(
    strat_birth_yr_childhood$tbls,
    strat_ADI_childhood$tbls,
    strat_edu_childhood$tbls,
    strat_income_childhood$tbls
  ),
  tab_spanner = tab_spanner
) |>  polish_childhood() |>
  modify_column_hide(
    c(
      exposure_1,
      exposure_2,
      exposure_3,
      exposure_4,
      exposure_5,
      exposure_6,
      exposure_7,
      exposure_8,
      exposure_9,
      exposure_10,
      exposure_11,
      exposure_12,
      exposure_13
    )
  )

strat_all_childhood
```

### Table 3: Association between childhood and adolescent factors and thyroid cancer incidence in the Sister Study participants, by DTC histology, tumor size, and cancer stage

```{r}
#| label: tbl-childhood-censor
#| tbl-cap: "Association between childhood and adolescent factors and thyroid cancer incidence in the Sister Study participants, by DTC histology, tumor size, and cancer stage"
#| cache: true
#| dependson: c("setup", "var-list-childhood", "functions")
#| cache.extra: !expr file.mtime("C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/data/ss_earlylife_popu.csv")

cox_PTC_childhood = cox_model(
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC_PTC)",
  var_name = var_name_childhood,
  var_label = var_label_childhood,
  covar = covariates_childhood,
  data = earlylife_popu
)

cox_stage_I_childhood = cox_model(
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC_stageI)",
  var_name = var_name_childhood,
  var_label = var_label_childhood,
  covar = covariates_childhood,
  data = earlylife_popu
)
cox_adv_stage_childhood = cox_model(
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC_adv_stage)",
  var_name = var_name_childhood,
  var_label = var_label_childhood,
  covar = covariates_childhood,
  data = earlylife_popu
)

cox_1cm_childhood = cox_model(
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC_1cm)",
  var_name = var_name_childhood,
  var_label = var_label_childhood,
  covar = covariates_childhood,
  data = earlylife_popu
)

cox_large_tumor_childhood = cox_model(
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC_large_tumor)",
  var_name = var_name_childhood,
  var_label = var_label_childhood,
  covar = covariates_childhood,
  data = earlylife_popu
)

sensi_censor_childhood = tbl_merge(
  tbls = list(cox_PTC, cox_stage_I, cox_adv_stage, cox_1cm, cox_large_tumor),
  tab_spanner = c(
    "**PTC**",
    "**Stage I**",
    "**Advanced stage**",
    "**Tumor size <1 cm**",
    "**Larger tumor**"
  )
) |>  polish_utero() |>
  modify_column_hide(c(exposure_1, exposure_2, exposure_3, exposure_4, exposure_5))

sensi_censor_childhood
```
