---
title: In-utero and infancy factors and thyroid cancer risk in adult women in the Sister Study cohort
toc: true
toc-expand: true
number-sections: true
anchor-sections: true
format:
  html:
    grid:
      sidebar-width: 50px
      margin-width:  200px
      body-width:	   2500px
    embed-resources: true
    code-fold: true
    code-summary: "Show the code"
    code-tools: true
    number_sections: true
date: 'last-modified'
date-format: '[This version:] MMMM DD, YYYY'
execute: 
  warning: false
  output: true
  message: false
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r}
#Time the code running process
start_time <- Sys.time()
```

```{r}
#| label: setup-library
###Necessary R library and data loading
library(here)
library(tidyverse)
library(gtsummary)
library(survival)
library(cmprsk)
library(broom)
library(broom.helpers)
library(tidymodels)
library(purrr)
library(splines)
library(Gmisc, quietly = TRUE)
library(glue)
library(htmlTable)
library(grid)
library(magrittr)


directory="C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS"

source(paste(directory,'/Early-life-Thyroid-cancer-SS/analysis/R_created_functions.R',sep = ""),local = knitr::knit_global())
```

```{r}
#| label: data
earlylife_popu =
  read.csv(paste(directory,"/data/ss_earlylife_popu.csv",sep = ""),
    #because quarto cannot knit the output document if here::here("data/ss_earlylife_popu.csv") is put inside read.csv like this,
    header = TRUE,
    stringsAsFactors = TRUE,
    fileEncoding = "latin1"
  )

earlylife_popu=earlylife_popu|>
  mutate(ident_DTC_PTC=ifelse(ident_DTC_histo=="2) PTC",1,0),
         ident_DTC_stageI=ifelse(ident_DTC_stage_AJCC7=="2) I",1,ident_DTC),
         ident_DTC_adv_stage=ifelse(ident_DTC_stage_AJCC7=="2) I",0,ident_DTC),
         ident_DTC_1cm=ifelse(ident_DTC_size=="2) <1 cm",1,ident_DTC),
         ident_DTC_large_tumor=ifelse(ident_DTC_size=="2) <1 cm",0,ident_DTC),
         ses_ADI_baseline_cat_half=case_when(
                                            ses_ADI_baseline_cat=="1) Quartile 1" ~ "1) Less than 30",
                                            ses_ADI_baseline_cat=="2) Quartile 2" ~ "1) Less than 30",
                                            ses_ADI_baseline_cat=="3) Quartile 3" ~ "2) 30 or more",
                                            ses_ADI_baseline_cat=="4) Quartile 4" ~ "2) 30 or more",
                                            ses_ADI_baseline_cat=="Unknown" ~ "Unknown"),
         ses_ADI_baseline_cat_half = factor(ses_ADI_baseline_cat_half, levels=c("1) Less than 30","2) 30 or more","Unknown")))
levels(earlylife_popu$ui_birth_weight_cat)=c("2) Between 2500 and 3999 g","1) < 2500 g","3) 4000+ g","Unknown")



```

```{r, file='C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/Early-life-Thyroid-cancer-SS/analysis/R_created_functions.R'}
#| label: function
```

```{r}
#| label: var-list-utero
###Create lists of exposure and covariates to avoid repetition in the code

#Covariates to adjust in Cox models
covariates_label_utero = list(
  "ses_race"~"Race",
  "an_bmi_cat"~"BMI",
  "smok_status"~"Smoking status",
  "illness_thyroid_all"~"Benign thyroid disease",
  "ses_edu_level"~"Educational level",
  "ses_income"~"Household annual income",
  "ses_ADI_baseline_cat_half"~"Area depreviation index"
)
covariates_utero = substr(covariates_label_utero, 2, unlist(gregexpr("~", covariates_label_utero)) - 3)

#In-utero and infancy factors
var_label_utero = list(
  "ui_diab_all_mother"~"Pre-pregnancy or gestational diabetes",
  "ui_hypertension_all"~"Gestational hypertension or hypertension-related disorders",
###"ui_epilepsy_prior"~"Epilepsy",                                            #not included because of number of cases<5
  "ui_vomit"~"Morning sickness with vomiting",
###"ui_DES"~"DES use",                                                        #not included because of number of cases<5
  "ui_mother_smok"~"Smoking during pregnancy",
  "ui_hh_smok"~"Household secondhand smoking during pregnancy",
  "ui_mother_age_cat"~"Age at delivery",
  "ui_mother_age_cont"~"Age at delivery (year, continuous)",
  "ui_diab_father"~"Diabetes",
  "ui_father_smok"~"Smoking during 3 months before conception",
  "ui_father_age_cat"~"Age at delivery",
  "ui_father_age_cont"~"Age at delivery (year, continuous)",
  "ui_birth_weight_cat"~"Birth weight",
  "ui_birth_weight_cont"~"Birth weight (continuous)",
  "ui_maternal_1st_born"~"First born",
  "ui_preg_order_cat"~"Birth order",
  "ui_preg_order_cont"~"Birth order (continuous)",
  "ui_preg_last_cat"~"Time since last pregnancy of the participants' mother",
  "ui_preg_last_cont"~"Time since last pregnancy of the participants' mother (year, continuous)",
  "ui_multiple_birth"~"Multiple birth",
  "ui_early_gest"~"Gestational age at birth",
###"ui_early_gest_detail"~"Gestational age at birth",                         #for exploratory purposes only, number of cases<5
  "ui_breastfed"~"Ever breastfed"
###"ui_soy"~"Ever fed soy formula"                                            #not included because of number of cases<5)
)
var_name_utero = substr(var_label_utero, 2, unlist(gregexpr("~", var_label_utero)) - 3)

```

# Main Tables and Figures

## Figure 1: Flowchart of study population

```{r}
#| label: flowchart
#| out-width: "50%"

eligible <- boxGrob(glue("Female participants", "in the Sister Study cohort",
                         "n = 50884",
                         pop = txtInt(1032),
                         .sep = "\n"),y = unit(0.5, "npc"),
  x = unit(0.2, "npc"))
included <- boxGrob(glue("Study population",
                         "n = 47810",
                         incl = txtInt(122),
                         .sep = "\n"),y = unit(0.5, "npc"),
  x = unit(0.2, "npc"))

excluded <- boxGrob(glue("Excluded:",
                         " - Withdrawal from the Sister study: n = 5",
                         " - Self-reported chemotherapy or radiotherapy for cancer prior",
                         " to baseline: n = 55",
                         " - Self-reported total thyroidectomy prior to baseline: n = 103",
                         " - Any history of invassive cancer prior to baseline: n = 2911",
                         tot = 30,
                         uninterested = 12,
                         contra = 30 - 12,
                         .sep = "\n"),
                    just = "left")

grid.newpage()
vert <- spreadVertical(eligible = eligible,
                       included = included)

excluded <- moveBox(excluded,
                    x = 0.65,
                    y =coords(vert$included)$top + distance(vert$eligible, vert$included, half = TRUE, center = FALSE))

for (i in 1:(length(vert) - 1)) {
  connectGrob(vert[[i]], vert[[i + 1]], type = "vert") %>%
    print
}

connectGrob(vert$eligible, excluded, type = "L")

# Print boxes
print(vert)
print(excluded)
rm(eligible,excluded,included,vert)

```

## Table 1: Characteristics of the study population (n=47810)

```{r}
#| label: tbl-descript
#| tbl-cap: "Characteristics of the study population (n=47810)"
#| cache: true
#| dependson: c("data", "var-list-utero", "functions")
#| cache.extra: !expr file.mtime("C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/data/ss_earlylife_popu.csv")

table1_descript_var_label = list(
  "ident_ageexact_bl"~"Age at baseline",
  "ses_race"~"Race",
  "an_bmi_cat"~"BMI",
  "smok_status"~"Smoking status",
  "illness_thyroid_all"~"Benign thyroid disease",
  "ses_edu_level"~"Educational level",
  "ses_income"~"Household annual income",
  "ses_ADI_baseline_cat_half"~"Area depreviation index"
)
table1_descript_var_name = substr(table1_descript_var_label, 2, unlist(gregexpr("~", table1_descript_var_label)) - 3)

descript_covar = descript_table(var_name = table1_descript_var_name,
                                var_label = table1_descript_var_label,
                                data = earlylife_popu) |> sort_birth_weight()
rm(table1_descript_var_label,table1_descript_var_name)
descript_covar
```

## Table 2: Association between in-utero and infancy factors and thyroid cancer incidence in the Sister Study participants

```{r}
#| label: tbl-utero
#| tbl-cap: "Association between in-utero and infancy factors and thyroid cancer incidence"
#| cache: true
#| dependson: c("data", "var-list-utero", "functions")
#| cache.extra: !expr file.mtime("C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/data/ss_earlylife_popu.csv")

###To create Table Association between in-utero and infancy factors and thyroid cancer incidence



####Loop over in utero and infancy factors in cox models, both univariate and multivariate analyses
descript_utero = descript_table(var_name = var_name_utero, 
                                var_label = var_label_utero,
                                data = earlylife_popu)|> sort_birth_weight()

univariate_utero=cox_model(surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
                     var_name = var_name_utero,
                     var_label = var_label_utero,
                     covar = NULL,
                     data=earlylife_popu)|> sort_birth_weight()

multivariate_utero=cox_model(surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
                       var_name = var_name_utero,
                       var_label = var_label_utero,
                       covar = covariates_utero,
                       data=earlylife_popu)|> sort_birth_weight()

#assemble descriptive table and cox models' results
cox_utero = tbl_merge(
  tbls = list(descript_utero, univariate_utero, multivariate_utero),
  tab_spanner = c("**N (%)**",
                  "**Univariate models**",
                  "**Multivariate models**")) |> 
  polish_utero()

#return the table 
rm(descript_utero, univariate_utero, multivariate_utero)
cox_utero
```

## Table 3: Association between birth weight (continuous) and thyroid cancer incidence in the Sister Study participants by Area Deprivation Index, levels of education, and household income

```{r}
#| label: tbl-birthweight-strat
#| tbl-cap: "Association between birth weight (continuous) and thyroid cancer incidence in the Sister Study participants by Area Deprivation Index, levels of education, and household income"

strat_ADI_birthweight = stratification_long(
  strat_var = ses_ADI_baseline_cat_half,
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
  var_name = "ui_birth_weight_cont",
  var_label = list("ui_birth_weight_cont" ~ "Birth weight (continuous)"),
  covar = covariates_utero,
  data = earlylife_popu
)

strat_edu_birthweight = stratification_long(
  strat_var = ses_edu_level,
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
  var_name = "ui_birth_weight_cont",
  var_label = list("ui_birth_weight_cont" ~ "Birth weight (continuous)"),
  covar = covariates_utero,
  data = earlylife_popu
)

strat_income_birthweight = stratification_long(
  strat_var = ses_income,
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
  var_name = "ui_birth_weight_cont",
  var_label = list("ui_birth_weight_cont" ~ "Birth weight (continuous)"),
  covar = covariates_utero,
  data = earlylife_popu
)


strat_all_birthweight = tbl_stack(list(
  strat_ADI_birthweight,
  strat_edu_birthweight,
  strat_income_birthweight
))

strat_all_birthweight = strat_all_birthweight |> modify_table_body(
  ~ .x  |>
    dplyr::add_row(
      label = "Area Deprivation Index",
      .before = grep("1) Less than 30", strat_all_birthweight$table_body$label)
    )  |>
    dplyr::add_row(
      label = "Level of education",
      .before = grep(
        "1) Highschool or GED or less",
        strat_all_birthweight$table_body$label
      ) + 1
    )  |>
    dplyr::add_row(
      label = "Household income",
      .after = grep(
        "3) Bachelors degree or higher",
        strat_all_birthweight$table_body$label
      ) + 2
    )
) |>
  modify_table_styling(
    column = "label",
    rows = strat_all_birthweight$table_body$label %in% c(
      "Area Deprivation Index",
      "Level of education",
      "Household income"
    ),
    text_format = "bold"
  )   

strat_all_birthweight
```

## Table 4: Simultaneously adjusted association between thyroid cancer risk and pre-pregnancy or gestational diabetes, gestational hypertension or hypertension-related disorders, birth weight, and gestational age at birth

```{r}
#| label: tbl-hyper-diab-birthweight
#| tbl-cap: "Simultaneously adjusted association between thyroid cancer risk and pre-pregnancy or gestational diabetes, gestational hypertension or hypertension-related disorders, birth weight, and gestational age at birth"

surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)"
mutual_name_cat = c("ui_diab_all_mother",
                    "ui_hypertension_all",
                    "ui_early_gest",
                    "ui_birth_weight_cat")
mutual_label_cat = list(
  "ui_diab_all_mother" ~ "Pre-pregnancy or gestational diabetes",
  "ui_hypertension_all" ~ "Gestational hypertension or pre-eclampsia",
  "ui_early_gest" ~ "Gestational age at birth",
  "ui_birth_weight_cat" ~ "Birth weight"
)

mutual_name_cont = c("ui_diab_all_mother",
                     "ui_hypertension_all",
                     "ui_early_gest",
                     "ui_birth_weight_cont")
mutual_label_cont = list(
  "ui_diab_all_mother" ~ "Pre-pregnancy or gestational diabetes",
  "ui_hypertension_all" ~ "Gestational hypertension or pre-eclampsia",
  "ui_early_gest" ~ "Gestational age at birth",
  "ui_birth_weight_cont" ~ "Birth weight (continuous)"
)

hyper_diab_birthweight_mutual_cat = reformulate(c(mutual_name_cat, covariates_utero), response = surv) |>
  coxph(method = "breslow", data = earlylife_popu) |>
  tbl_regression(
    exponentiate = TRUE,
    conf.int = TRUE,
    conf.level = 0.95,
    add_estimate_to_reference_rows = TRUE,
    label = mutual_label_cat,
    include = mutual_name_cat
  )|> sort_birth_weight()

hyper_diab_birthweight_mutual_cont = reformulate(c(mutual_name_cont, covariates_utero), response = surv) |>
  coxph(method = "breslow", data = earlylife_popu) |>
  tbl_regression(
    exponentiate = TRUE,
    conf.int = TRUE,
    conf.level = 0.95,
    add_estimate_to_reference_rows = TRUE,
    label = list("ui_birth_weight_cont" ~ "Birth weight (continuous)"),
    include = "ui_birth_weight_cont"
  )

hyper_diab_birthweight_mutual_all = tbl_stack(list(
  hyper_diab_birthweight_mutual_cat,
  hyper_diab_birthweight_mutual_cont
)) |>
  modify_fmt_fun(estimate ~ function(x)
    ifelse(is.na(x), NA, "—"),
    rows = n_event == "0") |>
  modify_fmt_fun(ci ~ function(x)
    ifelse(is.na(x), NA, "—"),
    rows = n_event == "0") |>
  modify_column_hide(p.value) |>
  modify_table_styling(
    columns = c(estimate, ci),
    rows = p.value < 0.05,
    text_format = "bold"
  ) |>
  modify_header(n_event ~ "**Number of events**",
                exposure ~ "**Person-years**",
                estimate ~ "**HR**") |>
  modify_table_body( ~ .x |> dplyr::relocate(exposure, .before = estimate))|>
  modify_footnote(abbreviation = TRUE,
                         `estimate` ~ "Multivariate models were adjusted for self-identified race/ethnicity, baseline BMI, smoking status, 
                         personal history of benign thyroid disease, educational level, household annual income, and Area Deprivation Index"
  ) 
hyper_diab_birthweight_mutual_all
```

# Supplementary materials

## Supplementary table 1: ICD codes used to classify well-differentiated thyroid cancer

| Histological categories   | Codes                                  |
|---------------------------|----------------------------------------|
| Papillary thyroid cancer  | 8050, 8260, 8340-8344, 8350, 8450-8460 |
| Follicular thyroid cancer | 8290, 8330-8335                        |
| Unknown histology         | 8000, 8010, missing data               |

## Supplementary table 2: Association between in-utero and infancy factors and thyroid cancer incidence in the Sister Study participants, when considering only medically confirmed cases, papillary histology, competing risks, or when stratified by tumor size, cancer stage, and birth cohort

```{r}
#| label: utero-medically-confirmed
#| output: false
#| cache: true
#| dependson: c("data", "var-list-utero", "functions")
#| cache.extra: !expr file.mtime("C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/data/ss_earlylife_popu.csv")

cox_medical_confirmed = cox_model(
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC_source)",
  var_name = var_name_utero,
  var_label = var_label_utero,
  covar = covariates_utero,
  data = earlylife_popu
)|> sort_birth_weight()

```

```{r}
#| label: utero-PTC
#| output: false
#| cache: true
#| dependson: c("data", "var-list-utero", "functions")
#| cache.extra: !expr file.mtime("C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/data/ss_earlylife_popu.csv")

cox_PTC_utero = cox_model(
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC_PTC)",
  var_name = var_name_utero,
  var_label = var_label_utero,
  covar = covariates_utero,
  data = earlylife_popu
)|> sort_birth_weight()

```

```{r}
#| label: utero-cmprsk
#| output: false
##| cache: true
##| dependson: c("data", "var-list-utero", "functions")
##| cache.extra: !expr file.mtime("C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/data/ss_earlylife_popu.csv")
 earlylife_popu$ident_compete_factor = factor(earlylife_popu$ident_compete)
 
 cox_cmprsk_utero = competing_risk_model(
   surv = "Surv(ident_ageexact_bl, ident_EOF, ident_compete_factor)",
   var_name = var_name_utero,
   var_label = var_label_utero,
   covar = covariates_utero,
   data = earlylife_popu
 )|> sort_birth_weight()

```

```{r}
#| label: utero-stageI
#| output: false
#| cache: true
#| dependson: c("data", "var-list-utero", "functions")
#| cache.extra: !expr file.mtime("C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/data/ss_earlylife_popu.csv")

cox_stage_I_utero = cox_model(
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC_stageI)",
  var_name = var_name_utero,
  var_label = var_label_utero,
  covar = covariates_utero,
  data = earlylife_popu
)|> sort_birth_weight()

```

```{r}
#| label: utero-advanced-stage
#| output: false
#| cache: true
#| dependson: c("data", "var-list-utero", "functions")
#| cache.extra: !expr file.mtime("C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/data/ss_earlylife_popu.csv")

cox_adv_stage_utero = cox_model(
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC_adv_stage)",
  var_name = var_name_utero,
  var_label = var_label_utero,
  covar = covariates_utero,
  data = earlylife_popu
)|> sort_birth_weight()

```

```{r}
#| label: utero-1cm
#| output: false
#| cache: true
#| dependson: c("data", "var-list-utero", "functions")
#| cache.extra: !expr file.mtime("C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/data/ss_earlylife_popu.csv")

cox_1cm_utero = cox_model(
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC_1cm)",
  var_name = var_name_utero,
  var_label = var_label_utero,
  covar = covariates_utero,
  data = earlylife_popu
)|> sort_birth_weight()

```

```{r}
#| label: utero-larger-tumor
#| output: false
#| cache: true
#| dependson: c("data", "var-list-utero", "functions")
#| cache.extra: !expr file.mtime("C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/data/ss_earlylife_popu.csv")

cox_large_tumor_utero = cox_model(
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC_large_tumor)",
  var_name = var_name_utero,
  var_label = var_label_utero,
  covar = covariates_utero,
  data = earlylife_popu
)|> sort_birth_weight()

```

```{r}
#| label: utero-strat-by-birth-cohort
#| cache: true
#| dependson: c("data", "var-list-utero", "functions")
#| cache.extra: !expr file.mtime("C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/data/ss_earlylife_popu.csv")

cox_birth_yr_utero = stratification_wide(
  strat_var = ident_birth_year,
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
  var_name = var_name_utero,
  var_label = var_label_utero,
  covar = covariates_utero,
  data = earlylife_popu
) |> sort_birth_weight()

cox_birth_yr_utero$table_styling$header$spanning_header =
  ifelse(
    cox_birth_yr_utero$table_styling$header$spanning_header == "**Before 1951**",
    "**Birth cohort: Before 1951**",
    ifelse(
      cox_birth_yr_utero$table_styling$header$spanning_header == "**1951 or after**",
      "**Birth cohort: 1951 or after**",
      cox_birth_yr_utero$table_styling$header$spanning_header
    )
  )

```

```{r}
#| label: tbl-utero-censor-all
#| tbl-cap: "Association between in-utero and infancy factors and thyroid cancer incidence in the Sister Study participants, when considering only medically confirmed cases, papillary histology, competing risks, or when stratified by tumor size, and cancer stage"
#| cache: true
#| dependson: c("utero-medically-confirmed","utero-PTC","utero-cmprsk","utero-stageI","utero-advanced-stage","utero-1cm","utero-larger-tumor")
#| cache.extra: !expr file.mtime("C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/data/ss_earlylife_popu.csv")

sensi_censor_utero = tbl_merge(
  tbls = list(cox_cmprsk_utero, cox_medical_confirmed, cox_PTC_utero,  
              cox_stage_I_utero, cox_adv_stage_utero, cox_1cm_utero, cox_large_tumor_utero,cox_birth_yr_utero),
  tab_spanner = c(
    "**Competing risks**",
    "**Medically confirmed DTC cases**",
    "**Papillary thyroid cancer**",
    "**Stage I thyroid cancer**",
    "**Advanced stage thyroid cancer**",
    "**Tumor size of less than 1 cm**",
    "**Tumor size of 1 cm or more**",
    "**Birth cohort: Before 1951**"
  )
) |>  polish_utero() |>
  modify_column_hide(c(exposure_1, exposure_2, exposure_3, exposure_4, exposure_5,exposure_6, exposure_7,exposure_1_8, exposure_2_8))

sensi_censor_utero$table_styling$header$spanning_header[213:238]=rep("**Birth cohort: 1951 or after**",26)
sensi_censor_utero
```

## Supplementary table 3: Association between birth weight and thyroid cancer risk, after excluding participants who were born more than 2 weeks before due date or those who were part of a multiple birth

```{r}
#| label: tbl-sensi-birthweight
#| tbl-cap: "Association between birth weight and thyroid cancer risk, after excluding participants who were born more than 2 weeks before due date or those who were part of a multiple birth"
surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)"

non_preterm = earlylife_popu |> filter(ui_early_gest != "2) Yes, at least 2 weeks before due date" & (ui_multiple_birth != "2) Multiple birth" ))


birth_weight_name = c("ui_birth_weight_cat", "ui_birth_weight_cont")
birth_weight_label = list(
  "ui_birth_weight_cat" ~ "Birth weight",
  "ui_birth_weight_cont" ~ "Birth weight (continuous)"
)

descript_birth = descript_table(var_name = birth_weight_name,
                                var_label = birth_weight_label,
                                data = non_preterm)

multivariate_birth = cox_model(
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
  var_name = birth_weight_name,
  var_label = birth_weight_label,
  covar = covariates_utero,
  data = non_preterm
)

#assemble descriptive table and cox models' results
cox_birth_weight = tbl_merge(
  tbls = list(descript_birth, multivariate_birth),
  tab_spanner = c("**N (%)**",
                  "**Multivariate models**")
) |> sort_birth_weight() |>
  #polish the table
  modify_footnote(
    abbreviation = TRUE,
    `estimate_2` ~ "Multivariate models were adjusted for self-identified race/ethnicity, baseline BMI, smoking status, 
                         personal history of benign thyroid disease, educational level, household annual income, and Area Deprivation Index"
  )
cox_birth_weight

```

# Nitty-gritty analyses that will not be shown in the paper

## Test for non-linearity of birth weight in the whole population and in non-preterm participants

Whole population, multivariate model, add squared term and compare the two nested models

```{r}
# earlylife_popu=earlylife_popu|>mutate(ui_birth_weight_cont_2= (ui_birth_weight_cont)^2)
# a = reformulate(c(covariates_utero, "ui_birth_weight_cont"), response = surv) |>
#   coxph(method = "breslow", data = earlylife_popu)
# b = reformulate(c(covariates_utero, "ui_birth_weight_cont", "ui_birth_weight_cont_2"),  response = surv) |>
#   coxph(method = "breslow", data = earlylife_popu) 
# anova(a,b)

```

Non-preterm, multivariate model, add squared term and compare the two nested models

```{r}
#Add ui_birth_weight_cont squared, insignificant
# non_preterm=non_preterm|>mutate(ui_birth_weight_cont_2= (ui_birth_weight_cont)^2)
# a = reformulate(c(covariates_utero, "ui_birth_weight_cont"), response = surv) |>
#   coxph(method = "breslow", data = non_preterm)
# b = reformulate(c(covariates_utero, "ui_birth_weight_cont", "ui_birth_weight_cont_2"),  response = surv) |>
#   coxph(method = "breslow", data = non_preterm) 
# anova(a,b)

```

Whole population, multivariate model, insignificant spline birth weight,

`{r} #Insignificant B-Spline surv="Surv(ident_ageexact_bl, ident_EOF, ident_DTC)"  reformulate(c(covariates_utero,"bs(ui_birth_weight_cont)"),response = surv)|>  coxph(method = "breslow", data = earlylife_popu) |> summary()}`

```{r}
stop_time=Sys.time() - start_time
print("Time to run the analyses: ")
print(stop_time)
```
