---
title: Childhood and adolescent factors and thyroid cancer risk in adult women in the Sister Study cohort
toc: true
toc-expand: true
number-sections: true
anchor-sections: true
format:
  html:
    grid:
      sidebar-width: 50px
      margin-width:  200px
      body-width:	   2500px
    embed-resources: true
    code-fold: true
    code-summary: "Show the code"
    code-tools: true
    number_sections: true
date: 'last-modified'
date-format: '[This version:] MMMM DD, YYYY'
execute: 
  warning: false
  output: true
  message: false
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r}
#Time the code running process
start_time <- Sys.time()
```

```{r}
#| label: function
source(paste(here::here(),'/analysis/R_created_functions.R',sep = ""),local = knitr::knit_global())
```

```{r}
#| label: setup-library-and-data
source(paste(here::here(),'/analysis/setup_library_and_data.R',sep = ""),local = knitr::knit_global())
```

# **Paper 2: Association between childhood and adolescent factors and thyroid cancer incidence**

```{r}

#| label: var-list-childhood
###Create lists of exposure and covariates to avoid repetition in the code

#Covariates to adjust in Cox models
covariates_label_childhood = list("ses_race"~"Race")
covariates_childhood = substr(covariates_label_childhood, 2, unlist(gregexpr("~", covariates_label_childhood)) - 3)

#Childhood factors
var_label_childhood = list(
  "an_weight_age10" ~ "Weight at age 10 vs. peers"    ,
  "an_height_age10" ~ "Height at age 10 vs. peers"    ,
  "an_weight_teen" ~ "Teenage weight vs. peers"      ,
  "an_age_tallest" ~ "Age reached full adult height (continuous)"      ,
  "rep_br_dev_age_cat" ~ "Age at breast development"  ,
  "rep_br_dev_age_cont" ~ "Age at breast development (continuous)" ,
  "rep_age_menarche_cat" ~ "Age at menarche"   ,
  "rep_age_menarche_cont" ~ "Age at menarche (continuous)"   ,
  #"rep_bc_ever" ~ "Ever used hormonal birth control" ,
  #"rep_bc_ado" ~ "Hormonal birth control under age 20"  ,
  "rep_bc_min_age_cat" ~ "Start age for hormonal birth control" ,
  #"rep_bc_min_age_cont" ~ "Start age for hormonal birth control (continuous, among women who used hormonal birth control)"    ,
  #"rep_bc_ado_dur_cat" ~ "Hormonal birth control duration under age 20",
  #"rep_bc_ado_dur_cont" ~ "Hormonal birth control duration under age 20 (continuous)"   ,
  #"ph_hrs_wk_before20_cat" ~ "Physical activity before age 20 (total hours/week)"    ,
  #"ph_hrs_wk_before20_cont" ~ "Physical activity before age 20 (total hours/week, continuous)"    ,
  "ph_MET_wk_before20_cat" ~ "Physical activity before age 20 (MET hours/week)"      ,
  "ph_MET_wk_before20_cont" ~ "Physical activity before age 20 (MET hours/week)" ,
  "alc_age_start_reg_cat" ~ "Age started drinking regularly"         ,
  "alc_drinks_before20_cat" ~ "Alcohol consumption under 20 (drinks/year)"     ,
  #"alc_drinks_before20_cont" ~ "Alcohol consumption under 20 (drinks/year, among women who drank before 20)"     ,
  # "alc_binge_time_519_cat" ~ "Times per year drank 4+ drinks under 20"     ,
  # "alc_binge_time_519_cont" ~ "Times per year drank 4+ drinks under 20"     ,
  #"alc_age_start_reg_cont" ~ "Age started drinking regularly"        ,
  "sm_age_start_smok_cat" ~ "Age started smoking"   ,
  #"sm_age_start_smok_cont" ~ "Age started smoking (continuous, among women who smoked)"  ,
  #"sm_ETS_before18" ~ "Secondhand smoking under 18"     ,
  #"smok_before20" ~ "Smoking before 20"       ,
  # "sm_pack_yr_1019_cat" ~ "Smoking from 10 and 19 (Pack-years)"    ,
  # "sm_pack_yr_1019_cont" ~ "Smoking from 10 and 19 (Pack-years)"  ,
  "sm_pack_yr_before20_cat" ~ "Smoking before 20 (Pack-years)"  ,
  #"sm_pack_yr_before20_cont" ~ "Smoking before 20 (Pack-years) (continuous)"  ,
  # "sm_yr_smok_1019_cat" ~ "Total years of smoking from 10 to 19"    ,
  # "sm_yr_smok_1019_cont" ~ "Total years of smoking from 10 to 19"   ,
  #"sm_yr_smok_before20_cat" ~ "Total years of smoking before 20"     ,
  #"sm_yr_smok_before20_cont" ~ "Total years of smoking before 20 (continuous)"   ,
   #"sm_age_stop_smok_cat" ~ "Age stopped smoking"    ,
   #"sm_age_stop_smok_cont" ~ "Age stopped smoking (continuous, among women who smoked)"   ,
  "sm_ETS_before18_yr_cat" ~ "Total years of secondhand smoking under 18"   ,
  #"sm_ETS_before18_yr_cont" ~ "Total years of secondhand smoking under 18 (continuous)"  ,
  "ses_income_child" ~ "Family income while growing up"         ,
  "ses_hunger_child" ~ "Ever not having enough to eat during childhood"          ,
  "ses_edu_level_child" ~ "Highest household education level at age 13",
  "ses_hh_compo_child" ~ "Household composition at age 13",
"ses_urban_child" ~ "Childhood residence"
)
var_name_childhood = substr(var_label_childhood, 2, unlist(gregexpr("~", var_label_childhood)) - 3)
```

# Main Tables and Figures

## Figure 1: Flowchart of study population

```{r}
#| label: flowchart
#| out-width: "50%"

eligible <- boxGrob(glue("Female participants", "in the Sister Study cohort",
                         "n = 50884",
                         pop = txtInt(1032),
                         .sep = "\n"),y = unit(0.5, "npc"),
  x = unit(0.2, "npc"))
included <- boxGrob(glue("Study population",
                         "n = 47913",
                         incl = txtInt(122),
                         .sep = "\n"),y = unit(0.5, "npc"),
  x = unit(0.2, "npc"))

excluded <- boxGrob(glue("Excluded:",
                         " - Any history of invasive cancer prior to baseline: n = 2911",
                         " - Self-reported chemotherapy or radiotherapy for cancer prior",
                         " to baseline: n = 55",
                         " - Withdrawal from the Sister Study: n = 5",
                         tot = 30,
                         uninterested = 12,
                         contra = 30 - 12,
                         .sep = "\n"),
                    just = "left")

grid.newpage()
vert <- spreadVertical(eligible = eligible,
                       included = included)

excluded <- moveBox(excluded,
                    x = 0.65,
                    y =coords(vert$included)$top + distance(vert$eligible, vert$included, half = FALSE, center = FALSE))

for (i in 1:(length(vert) - 1)) {
  connectGrob(vert[[i]], vert[[i + 1]], type = "vert") %>%
    print
}

connectGrob(vert$eligible, excluded, type = "L")

# Print boxes
print(vert)
print(excluded)
rm(eligible,excluded,included,vert)

```

## Table 1: Characteristics of the study population (n=47913)

```{r}
#| label: tbl-descript
#| tbl-cap: "Characteristics of the study population (n=47913)"

table1_descript_var_label = list(
  "ident_ageexact_bl"~"Age at baseline",
  "ses_race"~"Race",
  "an_bmi_cat"~"BMI",
  "illness_thyroid_all"~"Benign thyroid disease",
  "ses_edu_level"~"Educational level",
  "ses_income"~"Household annual income"
)
table1_descript_var_name = substr(table1_descript_var_label, 2, unlist(gregexpr("~", table1_descript_var_label)) - 3)

descript_covar = descript_table(var_name = table1_descript_var_name,
                                var_label = table1_descript_var_label,
                                data = earlylife_popu) |> sort_birth_weight()
rm(table1_descript_var_label,table1_descript_var_name)
descript_covar
```

## Table 2: Association between childhood and adolescent factors and thyroid cancer incidence in the Sister Study participants

```{r}
#| label: tbl-childhood
#| tbl-cap: "Association between childhood and adolescent factors and thyroid cancer incidence"

###To create Table Association between childhood and adolescent factors and thyroid cancer incidence

####Loop over childhood factors in cox models, both univariable and multivariable analyses
descript_childhood = descript_table(var_name = var_name_childhood, 
                                var_label = var_label_childhood,
                                data = earlylife_popu)|>sort_childhood()

multivariable_childhood=cox_model(surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
                             var_name = var_name_childhood,
                             var_label = var_label_childhood,
                             covar = covariates_childhood,
                             data=earlylife_popu)|>sort_childhood()

#assemble descriptive table and cox models' results
cox_childhood = tbl_merge(
  tbls = list(descript_childhood, multivariable_childhood),
  tab_spanner = c("**N (%)**","**Multivariable models**")) |> polish_childhood(group_name = TRUE)

#return the table 
cox_childhood
```

## Table 3: Simultaneously adjusted association between thyroid cancer risk and childhood growth and socioeconomic factors

```{r}
#| label: tbl-all-SES
#| tbl-cap: "Simultaneously adjusted association between thyroid cancer risk and childhood socioeconomic factors"

surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)"
mutual_label = list(
  "an_height_age10" ~ "Height at age 10 vs. peers",
  "an_weight_teen" ~ "Teenage weight vs. peers",
  "ses_hunger_child" ~ "Ever not having enough to eat during childhood"          ,
  "ses_edu_level_child" ~ "Highest household educational level at age 13",
  "ses_income_child" ~ "Family income while growing up"         ,
  "ses_hh_compo_child" ~ "Household composition at age 13",
  "ses_urban_child" ~ "Childhood residence"
  )
mutual_name = substr(mutual_label, 2, unlist(gregexpr("~", mutual_label)) - 3)


growth_SES_mutual = reformulate(c(mutual_name[1:4], covariates_childhood), response = surv) |>
  coxph(method = "breslow", data = earlylife_popu) |>
  tbl_regression(
    exponentiate = TRUE,
    conf.int = TRUE,
    conf.level = 0.95,
    add_estimate_to_reference_rows = TRUE,
    label = mutual_label[1:4],
    include = mutual_name[1:4]
  )|>modify_header(n_event ~ "**Number of events**")

growth_SES_mutual_all = reformulate(c(mutual_name, covariates_childhood), response = surv) |>
  coxph(method = "breslow", data = earlylife_popu) |>
  tbl_regression(
    exponentiate = TRUE,
    conf.int = TRUE,
    conf.level = 0.95,
    add_estimate_to_reference_rows = TRUE,
    label = mutual_label,
    include = mutual_name
  )

mutual=tbl_merge(list(growth_SES_mutual,growth_SES_mutual_all),tab_spanner = c("**Model 1**", "**Model 2**"))|> 
  sort_childhood() |>
  modify_fmt_fun(estimate_1 ~ function(x)
    ifelse(is.na(x), NA, "—"),
    rows = n_event_1 == "0") |>
  modify_fmt_fun(ci_1 ~ function(x)
    ifelse(is.na(x), NA, "—"),
    rows = n_event_1 == "0") |>
  modify_column_hide(p.value_1) |>
  modify_fmt_fun(estimate_2 ~ function(x)
    ifelse(is.na(x), NA, "—"),
    rows = n_event_2 == "0") |>
  modify_fmt_fun(ci_2 ~ function(x)
    ifelse(is.na(x), NA, "—"),
    rows = n_event_2 == "0") |>
  modify_column_hide(p.value_2) |>
  modify_footnote(abbreviation = TRUE,
                         `estimate_1` ~ "Model 1 was simultaneously adjusted for attained age (timescale), self-identified race/ethnicity, height at age 10 vs. peers, and teenage weight vs. peers, ever not having enough to eat during childhood, and highest household educational level at age 13. Model 2 was simultaneously adjusted for all covariates of model 1 and family income while growing up, household composition at age 13, and childhood residence") 
mutual
```

## Figure 3: The joint associations of teenage weight vs. peers and baseline BMI with differentiated thyroid cancer incidence and of weight at age 10 vs. peers and baseline BMI with differentiated thyroid cancer incidence

```{r}
#| label: joint-weight-figure
###Create lists of exposure and covariates that have a significant association

earlylife_popu = earlylife_popu |>
  mutate(ado_weight_adult_bmi = as.factor(paste0(an_weight_teen, " than peers, baseline BMI ", an_bmi_cat)))|>
  mutate(ado_weight_adult_bmi = fct_relevel(ado_weight_adult_bmi,c("Same weight than peers, baseline BMI <25", 
                                                                   "Same weight than peers, baseline BMI 25.0-29.9",     
                                                                   "Same weight than peers, baseline BMI 30.0+",
                                                                   "Lighter than peers, baseline BMI <25",
                                                                   "Lighter than peers, baseline BMI 25.0-29.9", 
                                                                   "Lighter than peers, baseline BMI 30.0+",
                                                                   "Heavier than peers, baseline BMI <25", 
                                                                   "Heavier than peers, baseline BMI 25.0-29.9",
                                                                   "Heavier than peers, baseline BMI 30.0+",
                                                                   "Heavier than peers, baseline BMI Unknown",
                                                                   "Lighter than peers, baseline BMI Unknown",
                                                                   "Same weight than peers, baseline BMI Unknown",  
                                                                   "Unknown than peers, baseline BMI <25", 
                                                                   "Unknown than peers, baseline BMI 25.0-29.9",
                                                                   "Unknown than peers, baseline BMI 30.0+")))|>
  arrange(ado_weight_adult_bmi)|>
  mutate(child_weight_adult_bmi = as.factor(paste0(an_weight_age10, " than peers, baseline BMI ", an_bmi_cat)))|>
  mutate(child_weight_adult_bmi = fct_relevel(child_weight_adult_bmi,c("Same weight than peers, baseline BMI <25", 
                                                                   "Same weight than peers, baseline BMI 25.0-29.9",     
                                                                   "Same weight than peers, baseline BMI 30.0+",
                                                                   "Lighter than peers, baseline BMI <25",
                                                                   "Lighter than peers, baseline BMI 25.0-29.9", 
                                                                   "Lighter than peers, baseline BMI 30.0+",
                                                                   "Heavier than peers, baseline BMI <25", 
                                                                   "Heavier than peers, baseline BMI 25.0-29.9",
                                                                   "Heavier than peers, baseline BMI 30.0+",
                                                                   "Heavier than peers, baseline BMI Unknown",
                                                                   "Lighter than peers, baseline BMI Unknown",
                                                                   "Same weight than peers, baseline BMI Unknown",  
                                                                   "Unknown than peers, baseline BMI <25", 
                                                                   "Unknown than peers, baseline BMI 25.0-29.9",
                                                                   "Unknown than peers, baseline BMI 30.0+")))|>
  arrange(child_weight_adult_bmi)

ado_weight_adult_bmi=cox_model(surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
          var_name = "ado_weight_adult_bmi",
          var_label = list("ado_weight_adult_bmi"~"Teenage weight vs. peers and baseline BMI"),
          covar = covariates_childhood,
          data=earlylife_popu)|>sort_childhood()

child_weight_adult_bmi=cox_model(surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
                               var_name = "child_weight_adult_bmi",
                               var_label = list("child_weight_adult_bmi"~"Weight at age 10 vs. peers and baseline BMI"),
                               covar = covariates_childhood,
                               data=earlylife_popu)|>sort_childhood()

results_ado_weight_adult_bmi=ado_weight_adult_bmi$table_body|>filter(n_event>0)|>
  mutate(group=c("Same weight (ref)","Same weight","Same weight",rep("Lighter",3),rep("Heavier",3)))|>
  mutate(group=fct_relevel(group,c("Lighter","Same weight","Same weight (ref)","Heavier")))|>
  mutate(bmi=as_factor(rep(c("BMI <25","BMI 25.0-29.9", "BMI 30.0+"),3)))

results_child_weight_adult_bmi=child_weight_adult_bmi$table_body|>filter(n_event>1)|> #n_event>1 to exclude Unknown category
  mutate(group=c("Same weight (ref)","Same weight","Same weight",rep("Lighter",3),rep("Heavier",3)))|>
  mutate(group=fct_relevel(group,c("Lighter","Same weight","Same weight (ref)","Heavier")))|>
  mutate(bmi=as_factor(rep(c("BMI <25","BMI 25.0-29.9", "BMI 30.0+"),3)))


results_ado_weight_adult_bmi = results_ado_weight_adult_bmi|>
  ggplot(aes(x = group,y = estimate, ymin = conf.low, ymax = conf.high ))+
  geom_pointrange(cex=3)+
  geom_hline(aes(fill=bmi),yintercept =1, linetype=2)+
  xlab('Teenage weight vs. peers')+ ylab("Hazard Ratio (95% Confidence Interval)")+
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high),width=0,cex=1.5)+ 
  facet_wrap(~bmi,strip.position="bottom",nrow=1,scales = "free_x") +
  #labs(tag = "B") +
  theme(plot.title=element_text(size=32,face="bold"),
        axis.text.x=element_text(size=32,face="bold",angle = 60, vjust = 1, hjust=1,margin = margin(t = 0, r = 10, b = 10, l = 0)),
        axis.text.y=element_text(size=32,face="bold",angle = 00, vjust = 1, hjust=1,margin = margin(t = 0, r = 10, b = 10, l = 0)),
        axis.title.x=element_text(size=32,face="bold",vjust = 3, hjust=0.5,margin = margin(t = 30, r = 0, b = 0, l = 0)),
        axis.title.y=element_text(size=32,face="bold",vjust = 3, hjust=0.5,margin = margin(t = 0, r = 10, b = 0, l = 20)),
        strip.text = element_text(size = 32, vjust = 1, hjust=0.5),
        aspect.ratio = 3.5/1.5,    strip.placement = "outside",
        panel.spacing = unit(2, "lines"),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        plot.margin=unit(c(3,3,3,3),"cm")
        #plot.tag = element_text(size=40,face="bold",margin = margin(t = 0, r = 20, b = 0, l = 0))
  )+
  scale_y_log10(breaks=c(0.5,1,2,3,4,5))+ylim(0.3,5.7)


results_child_weight_adult_bmi=results_child_weight_adult_bmi|>
  ggplot(aes(x = group,y = estimate, ymin = conf.low, ymax = conf.high ))+
  geom_pointrange(cex=3)+
  geom_hline(aes(fill=bmi),yintercept =1, linetype=2)+
  xlab('Weight at age 10 vs. peers')+ ylab("Hazard Ratio (95% Confidence Interval)")+
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high),width=0,cex=1.5)+ 
  #labs(tag = "A") +
  facet_wrap(~bmi,strip.position="bottom",nrow=1,scales = "free_x") +
  theme(plot.title=element_text(size=32,face="bold"),
        axis.text.x=element_text(size=32,face="bold",angle = 60, vjust = 1, hjust=1,margin = margin(t = 0, r = 10, b = 10, l = 0)),
        axis.text.y=element_text(size=32,face="bold",angle = 00, vjust = 1, hjust=1,margin = margin(t = 0, r = 10, b = 10, l = 0)),
        axis.title.x=element_text(size=32,face="bold",vjust = 3, hjust=0.5,margin = margin(t = 30, r = 0, b = 0, l = 0)),
        axis.title.y=element_text(size=32,face="bold",vjust = 3, hjust=0.5,margin = margin(t = 0, r = 10, b = 0, l = 20)),
        strip.text = element_text(size = 32, vjust = 1, hjust=0.5),
        aspect.ratio = 3.5/1.5,    strip.placement = "outside",
        panel.spacing = unit(2, "lines"),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        plot.margin=unit(c(3,3,3,3),"cm")
        #plot.tag = element_text(size=40,face="bold",margin = margin(t = 0, r = 20, b = 0, l = 0))
  )+
  scale_y_log10(breaks=c(0.5,1,2,3,4,5))+ylim(0.3,5.7)

png(filename=here::here("child_teen_bmi_weight_strat.png"),pointsize=14,width = 2400, height = 1500)
grid.arrange(results_child_weight_adult_bmi, results_ado_weight_adult_bmi, ncol=2)
dev.off()

```

## Supplementary materials

## Supplementary Figure 1: Association between height at age 10 vs peers, teenage weight vs peers, ever not having enough to eat during childhood, and highest household education level at age 13 and thyroid cancer incidence in the Sister Study participants stratified according to potential modifying factors

```{r}
#| label: childhood-strat-by-adult_SES
#| cache: true
ado_weight_adult_bmi_no_interaction=coxph(Surv(ident_ageexact_bl, ident_EOF, ident_DTC)~an_weight_teen+an_bmi_cat+ses_race+smok_status+illness_thyroid_all+
                                            ses_edu_level+ses_income+ses_ADI_baseline_cat,
                                          method = "breslow", data = earlylife_popu)
ado_weight_adult_bmi_with_interaction=coxph(Surv(ident_ageexact_bl, ident_EOF, ident_DTC)~an_weight_teen*an_bmi_cat+ses_race+smok_status+illness_thyroid_all+
                                            ses_edu_level+ses_income+ses_ADI_baseline_cat,
                                          method = "breslow", data = earlylife_popu)
anova(ado_weight_adult_bmi_no_interaction,ado_weight_adult_bmi_with_interaction)

strat_list = list(
  "an_bmi_cat" ~ "BMI",
  "ses_edu_level" ~ "Personal educational level",
  "ses_income" ~ "Household annual income")
strat_height_age10 = stratification_long_for_a_list(var_name = "an_height_age10",
                                                    var_label = list("an_height_age10" ~ "Height at age 10 vs. peers"),
                                                    strat_list =  strat_list,
                                                    surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
                                                    covar = covariates_childhood,
                                                    data = earlylife_popu) 
png(filename=here::here("childhood_height_age10_strat.png"),pointsize=14,width = 1000, height = 1200)
strat_height_age10 |>forest_plot_child(cat = "Taller",label1 = "Height at age 10 vs. peers: ",label2 = "Taller vs same height")
dev.off()

strat_weight_teen = stratification_long_for_a_list(var_name = "an_weight_teen",
                                                         var_label = list("an_weight_teen" ~ "Teenage weight vs. peers"),
                                                         strat_list =  strat_list,
                                                         surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
                                                         covar = covariates_childhood,
                                                         data = earlylife_popu) 
png(filename=here::here("childhood_weight_teen_lighter_strat.png"),pointsize=14,width = 1000, height = 1200)
strat_weight_teen|>forest_plot_child(cat = "Lighter",label1 = "Teenage weight vs. peers: ",label2 = "Lighter vs same weight")
dev.off()

strat_hunger_child = stratification_long_for_a_list(var_name = "ses_hunger_child",
                                                    var_label = list("ses_hunger_child" ~ "Ever not having enough to eat during childhood"),
                                                    strat_list =  strat_list,surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
                                                    covar = covariates_childhood,
                                                    data = earlylife_popu) 
png(filename=here::here("childhood_hunger_strat.png"),pointsize=14,width = 1000, height = 1200)
strat_hunger_child|>forest_plot_child(cat = "Yes",label1 = "Ever not having enough to eat",label2 = "during childhood: Yes vs No")
dev.off()

strat_edu_child = stratification_long_for_a_list(var_name = "ses_edu_level_child",
                                                 var_label = list("ses_edu_level_child" ~ "Highest household education level at age 13"),
                                                 strat_list =  strat_list,
                                                 surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
                                                 covar = covariates_childhood,
                                                 data = earlylife_popu) 
png(filename=here::here("childhood_edu_household_strat.png"),pointsize=14,width = 1200, height = 1200)
strat_edu_child|> forest_plot_child(cat = "Bachelor's degree or higher",label1 = "Highest household education level at age 13:",label2 = "Highschool or GED or less")
dev.off()

png(filename=here::here("childhood_edu_household_strat_2.png"),pointsize=14,width = 1200, height = 1200)
strat_edu_child|> forest_plot_child(cat = "Bachelor's degree or higher",label1 = "Bachelor's degree or higher vs",label2 = "Highschool or GED or less")
dev.off()
```

## Supplementary table 1: The joint associations of teenage weight vs. peers and baseline BMI with differentiated thyroid cancer incidence 

```{r}
#| label: joint-teenage-weight-table
ado_weight_adult_bmi
```

## Supplementary table 2: The joint associations of weight at age 10 vs. peers and baseline BMI with differentiated thyroid cancer incidence 

```{r}
#| label: joint-age10-weight-table
child_weight_adult_bmi
```


## Supplementary table 3: Association between childhood and adolescent factors and thyroid cancer incidence in the Sister Study participants, when considering only medically confirmed cases, papillary histology, competing risks, or when stratified by tumor size, cancer stage, and birth cohort

```{r}
#| label: var-list-childhood-for-sup-tables
###Create lists of exposure and covariates that have a significant association


var_label_childhood_sup = list(
  "an_height_age10" ~ "Height at age 10 vs. peers",
  "an_weight_teen" ~ "Teenage weight vs. peers",
  "ses_hunger_child" ~ "Ever not having enough to eat during childhood",
  "ses_edu_level_child" ~ "Highest household education level at age 13"
)
var_name_childhood_sup = substr(var_label_childhood_sup, 2, unlist(gregexpr("~", var_label_childhood_sup)) - 3)

```




```{r}
#| label: childhood-medically-confirmed
#| output: false

cox_medical_confirmed = cox_model(
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC_source)",
  var_name = var_name_childhood_sup,
  var_label = var_label_childhood_sup,
  covar = covariates_childhood,
  data = earlylife_popu
)|>sort_childhood()

```

```{r}
#| label: childhood-PTC
#| output: false

cox_PTC_childhood = cox_model(
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC_PTC)",
  var_name = var_name_childhood_sup,
  var_label = var_label_childhood_sup,
  covar = covariates_childhood,
  data = earlylife_popu
)|>sort_childhood()

```

```{r}
#| label: childhood-cmprsk
#| output: false
earlylife_popu$ident_compete_factor = factor(earlylife_popu$ident_compete)
  
cox_cmprsk_childhood = competing_risk_model(
    surv = "Surv(ident_ageexact_bl, ident_EOF, ident_compete_factor)",
    var_name = var_name_childhood_sup,
    var_label = var_label_childhood_sup,
    covar = covariates_childhood,
    data = earlylife_popu
  )|>sort_childhood()

```

```{r}
#| label: early-onset

earlylife_popu_early_onset=earlylife_popu|>filter(ident_ageexact_bl<50)|>
  mutate(ident_EOF_early_onset=ifelse(ident_EOF>50,50,ident_EOF),
         ident_DTC_early_onset=ifelse(ident_EOF>50,0,ident_DTC))

cox_early_onset_childhood = cox_model(
  surv = "Surv(ident_ageexact_bl, ident_EOF_early_onset, ident_DTC_early_onset)",
  var_name = var_name_childhood_sup,
  var_label = var_label_childhood_sup,
  covar = covariates_childhood,
  data = earlylife_popu_early_onset
)|>sort_childhood()

```

```{r}
#| label: late-onset

earlylife_popu_late_onset=earlylife_popu|>filter(ident_ageexact_bl>=50)
cox_late_onset_childhood = cox_model(
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC_PTC)",
  var_name = var_name_childhood_sup,
  var_label = var_label_childhood_sup,
  covar = covariates_childhood,
  data = earlylife_popu_late_onset
)|>sort_childhood()

```

```{r}
#| label: e-values
#| output: false
multivariable_evalues=multivariable_childhood|>modify_table_body(filter,variable %in% c("an_height_age10",   "an_weight_teen","ses_hunger_child","ses_edu_level_child"))|>polish_childhood(group_name = TRUE)
Evalues=cbind(multivariable_evalues$`_data`$estimate,multivariable_evalues$`_data`$conf.low,multivariable_evalues$`_data`$conf.high)

colnames(Evalues)=c("estimate","conf.low","conf.high")

Evalues=round(add_evalues(Evalues),2)
Evalues=cbind(multivariable_evalues$`_data`$label,Evalues)|>as.data.frame()
```

```{r}
#| label: tbl-childhood-censor-all
#| tbl-cap: "Association between childhood and adolescent factors and thyroid cancer incidence in the Sister Study participants, when considering only medically confirmed cases, papillary histology, competing risks, or when stratified by tumor size, and cancer stage"
##| cache: true
##| dependson: c("childhood-medically-confirmed","childhood-PTC","childhood-cmprsk","childhood-stageI","childhood-advanced-stage","childhood-1cm","childhood-larger-tumor")
##| cache.extra: !expr file.mtime("C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/data/ss_earlylife_popu.csv")

sensi_censor_childhood = tbl_merge(
  tbls = list(cox_early_onset_childhood,cox_late_onset_childhood, 
              cox_medical_confirmed, cox_PTC_childhood),
  tab_spanner = c(
    "**Early-onset cancer**",
    "**Late-onset cancer**",
    "**Medically confirmed DTC cases**",
    "**Papillary thyroid cancer**"
  )
) 
sensi_censor_childhood|> polish_childhood()
```

```{r}
#| label: tbl-childhood-censor-all
#| tbl-cap: "Association between childhood and adolescent factors and thyroid cancer incidence in the Sister Study participants, when considering only medically confirmed cases, papillary histology, competing risks, or when stratified by tumor size, and cancer stage"
sensi_censor_childhood
cox_cmprsk_childhood
Evalues
```
