---
title: Childhood and adolescent factors and thyroid cancer risk in adult women in the Sister Study cohort
toc: true
toc-expand: true
number-sections: true
anchor-sections: true
format:
  html:
    grid:
      sidebar-width: 50px
      margin-width:  200px
      body-width:	   2500px
    embed-resources: true
    code-fold: true
    code-summary: "Show the code"
    code-tools: true
    number_sections: true
date: 'last-modified'
date-format: '[This version:] MMMM DD, YYYY'
execute: 
  warning: false
  output: true
  message: false
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r}
#Time the code running process
start_time <- Sys.time()
```

```{r}
#| label: setup-library
###Necessary R library and data loading
library(here)
library(tidyverse)
library(gtsummary)
library(survival)
library(cmprsk)
library(broom)
library(broom.helpers)
library(tidymodels)
library(purrr)
library(splines)
library(Gmisc, quietly = TRUE)
library(glue)
library(htmlTable)
library(grid)
library(magrittr)


directory="C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS"

source(paste(directory,'/Early-life-Thyroid-cancer-SS/analysis/R_created_functions.R',sep = ""),local = knitr::knit_global())
```

```{r}
#| label: data
earlylife_popu =
  read.csv(paste(directory,"/data/ss_earlylife_popu.csv",sep = ""),
    #because quarto cannot knit the output document if here::here("data/ss_earlylife_popu.csv") is put inside read.csv like this,
    header = TRUE,
    stringsAsFactors = TRUE,
    fileEncoding = "latin1"
  )

earlylife_popu = earlylife_popu |>
  mutate(
    ident_DTC_PTC = ifelse(ident_DTC_histo == "2) PTC", 1, 0),
    ident_DTC_stageI = ifelse(ident_DTC_stage_AJCC7 == "2) I", 1, ident_DTC),
    ident_DTC_adv_stage = ifelse(ident_DTC_stage_AJCC7 == "2) I", 0, ident_DTC),
    ident_DTC_1cm = ifelse(ident_DTC_size == "2) <1 cm", 1, ident_DTC),
    ident_DTC_large_tumor = ifelse(ident_DTC_size == "2) <1 cm", 0, ident_DTC),
    ses_ADI_baseline_cat_half = case_when(
      ses_ADI_baseline_cat == "1) Quartile 1" ~ "1) Less than 30",
      ses_ADI_baseline_cat == "2) Quartile 2" ~ "1) Less than 30",
      ses_ADI_baseline_cat == "3) Quartile 3" ~ "2) 30 or more",
      ses_ADI_baseline_cat == "4) Quartile 4" ~ "2) 30 or more",
      ses_ADI_baseline_cat == "Unknown" ~ "Unknown"),
    ses_ADI_baseline_cat_half = factor(ses_ADI_baseline_cat_half, levels = c("1) Less than 30", "2) 30 or more", "Unknown")))

levels(earlylife_popu$ses_hh_compo_child)[levels(earlylife_popu$ses_hh_compo_child)=="3) Other"] = "Unknown"

levels(earlylife_popu$rep_br_dev_age_cat)=c("2) 11-13 years of age","1) Less than 11 years of age", "3) 14 years of age or more","Unknown")
levels(earlylife_popu$rep_age_menarche_cat)=c("2) 12-14 years of age","1) Less than 12 years of age","3) 14 years of age or more","Unknown")
levels(earlylife_popu$an_weight_age10)=c("2) Same weight","1) Lighter","3) Heavier","Unknown")
levels(earlylife_popu$an_height_age10)=c("2) Same height","1) Shorter","3) Taller","Unknown")
levels(earlylife_popu$an_weight_teen)=c("2) Same weight","1) Lighter","3) Heavier","Unknown")
levels(earlylife_popu$ses_income_child)=c("2) Middle income","1) Well off","3) Low income","4) Poor","5) Unknown")



```

```{r, file='C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/Early-life-Thyroid-cancer-SS/analysis/R_created_functions.R'}
#| label: function
```

# **Paper 2: Association between childhood and adolescent factors and thyroid cancer incidence**

```{r}

#| label: var-list-childhood
###Create lists of exposure and covariates to avoid repetition in the code

#Covariates to adjust in Cox models
covariates_label_childhood = list(
  "ses_race"~"Race",
  "an_bmi_cat"~"BMI",
  "smok_status"~"Smoking status",
  "illness_thyroid_all"~"Benign thyroid disease",
  "ses_edu_level"~"Educational level",
  "ses_income"~"Household annual income",
  "ses_ADI_baseline_cat_half"~"Area depreviation index"
)
covariates_childhood = substr(covariates_label_childhood, 2, unlist(gregexpr("~", covariates_label_childhood)) - 3)

# add_covariates_label = list(
#   "rep_meno_age_cat"~"Age at menopause"
#   "ph_MET_wk_current_cat"~"Physical activity (MET hours/week)"  ,
#   "ph_MET_wk_current_cont"~"Physical activity (MET hours/week)"  ,
#   "PH_CurrentTotMETHrsPerWeek_CDC"~"Physical activity (MET hours/week)",
#   "ph_hrs_wk_current_cont"~"Physical activity (total hours/week)"    ,
#   "ses_marital_status"~"Marital status"
#   "sm_pack_yr_cat"~"Smoking (pack/year)"            ,
#   "sm_pack_yr_cont"~"Smoking (pack/year)"
#   "sm_yr_smok_cat"~"Years of smoking"    ,
#   "sm_yr_smok_cont"~"Years of smoking"   ,
#   "sm_ETS_ever"~"Ever secondhand smoking"       ,
#   "sm_ETS_total_yr_cat"~"Total years of secondhand smoking"    ,
#   "sm_ETS_total_yr_cont"~"Total years of secondhand smoking"
# )
# add_covariates = substr(add_covariates_label, 2, unlist(gregexpr("~", add_covariates_label)) - 3)

#Childhood factors
var_label_childhood = list(
  "an_weight_age10" ~ "Weight at age 10 vs. peers"    ,
  "an_height_age10" ~ "Height at age 10 vs. peers"    ,
  "an_weight_teen" ~ "Teenage weight vs. peers"      ,
  "an_age_tallest" ~ "Age reached full adult height"      ,
  "rep_br_dev_age_cat" ~ "Age at breast development"  ,
  "rep_br_dev_age_cont" ~ "Age at breast development" ,
  "rep_age_menarche_cat" ~ "Age at menarche"   ,
  "rep_age_menarche_cont" ~ "Age at menarche"   ,
  "rep_bc_ever" ~ "Ever used hormonal birth control" ,
  #"rep_bc_ado" ~ "Hormonal birth control under age 20"  ,
  "rep_bc_ado_dur_cat" ~ "Hormonal birth control duration under age 20",
  "rep_bc_ado_dur_cont" ~ "Hormonal birth control duration under age 20"   ,
  "rep_bc_min_age_cat" ~ "Start age for hormonal birth control" ,
  "rep_bc_min_age_cont" ~ "Start age for hormonal birth control"    ,
  # "rep_proges_ado"~"Progestin-only birth control under age 20"     ,
  # "rep_proges_ado_dur_cat"~"Progestin-only birth control duration under age 20" ,
  # "rep_proges_ado_dur_cont"~"Progestin-only birth control duration under age 20"    ,
  # "rep_proges_min_age_cat"~"Start age for progestin-only birth control"    ,
  # "rep_proges_min_age_cont"~"Start age for progestin-only birth control"   ,
  # "rep_combined_ado"~"Combined hormonal birth control under age 20" ,
  # "rep_combined_ado_dur_cont"~"Combined hormonal birth control duration under age 20" ,
  # "rep_combined_ado_dur_cat"~"Combined hormonal birth control duration under age 20"   ,
  # "rep_combined_min_age_cont"~"Start age for combined hormonal birth control"   ,
  # "rep_combined_min_age_cat"~"Start age for combined hormonal birth control"  ,
  # "rep_bc_ado_type"~"Types of hormonal birth control under age 20"      ,
  # "rep_bc_ingredient"~"Ingredient of hormonal birth control under age 20"   ,
  # "ph_hrs_wk_child_cat" ~ "Physical activity before age 13 (total hours/week)"   ,
  # "ph_hrs_wk_child_cont" ~ "Physical activity before age 13 (total hours/week)"   ,
  # "ph_MET_wk_child_cat" ~ "Physical activity before age 13 (MET hours/week)"  ,
  # "ph_MET_wk_child_cont" ~ "Physical activity before age 13 (MET hours/week)"   ,
  # "ph_hrs_wk_teen_cat" ~ "Physical activity from 13 and 19 (total hours/week)"  ,
  # "ph_hrs_wk_teen_cont" ~ "Physical activity from 13 and 19 (total hours/week)"  ,
  # "ph_MET_wk_teen_cat" ~ "Physical activity between the age of 13 and 19 (MET hours/week)"    ,
  # "ph_MET_wk_teen_cont" ~ "Physical activity between the age of 13 and 19 (MET hours/week)"   ,
  #"ph_MET_wk_before20_cat" ~ "Physical activity before age 20 (MET hours/week)"      ,
  #"ph_MET_wk_before20_cont" ~ "Physical activity before age 20 (MET hours/week)" ,
  "ph_hrs_wk_before20_cat" ~ "Physical activity before age 20 (total hours/week)"    ,
  "ph_hrs_wk_before20_cont" ~ "Physical activity before age 20 (total hours/week)"    ,
  "alc_drinks_before20_cat" ~ "Alcohol consumption under 20 (drinks/year)"     ,
  #"alc_drinks_before20_cont" ~ "Alcohol consumption under 20 (drinks/year)"     ,
  # "alc_binge_time_519_cat" ~ "Times per year drank 4+ drinks under 20"     ,
  # "alc_binge_time_519_cont" ~ "Times per year drank 4+ drinks under 20"     ,
  "alc_age_start_reg_cat" ~ "Age started drinking regularly"         ,
  #"alc_age_start_reg_cont" ~ "Age started drinking regularly"        ,
  "sm_age_start_smok_cat" ~ "Age started smoking"   ,
  #"sm_age_start_smok_cont" ~ "Age started smoking"  ,
  # "sm_age_stop_smok_cat" ~ "Age stopped smoking"    ,
  # "sm_age_stop_smok_cont" ~ "Age stopped smoking"   ,
  #"sm_ETS_before18" ~ "Secondhand smoking under 18"     ,
  "sm_ETS_before18_yr_cat" ~ "Total years of secondhand smoking under 18"   ,
  "sm_ETS_before18_yr_cont" ~ "Total years of secondhand smoking under 18"  ,
  #"smok_before20" ~ "Smoking before 20"       ,
  # "sm_pack_yr_1019_cat" ~ "Smoking from 10 and 19 (Pack-years)"    ,
  # "sm_pack_yr_1019_cont" ~ "Smoking from 10 and 19 (Pack-years)"  ,
  "sm_pack_yr_before20_cat" ~ "Smoking before 20 (Pack-years)"  ,
  "sm_pack_yr_before20_cont" ~ "Smoking before 20 (Pack-years)"  ,
  # "sm_yr_smok_1019_cat" ~ "Total years of smoking from 10 to 19"    ,
  # "sm_yr_smok_1019_cont" ~ "Total years of smoking from 10 to 19"   ,
  "sm_yr_smok_before20_cat" ~ "Total years of smoking before 20"     ,
  "sm_yr_smok_before20_cont" ~ "Total years of smoking before 20"   ,
  "ses_income_child" ~ "Family income while growing up"         ,
  "ses_hunger_child" ~ "Ever not having enough to eat during childhood"          ,
  "ses_edu_level_child" ~ "Highest household education level at age 13",
  "ses_hh_compo_child" ~ "Household composition at age 13",
"ses_urban_child" ~ "Childhood residence"
)
var_name_childhood = substr(var_label_childhood, 2, unlist(gregexpr("~", var_label_childhood)) - 3)
smok=c("sm_age_start_smok_cat","sm_age_start_smok_cont")
```

# Main Tables and Figures

## Figure 1: Flowchart of study population

```{r}
#| label: flowchart
#| out-width: "50%"

eligible <- boxGrob(glue("Female participants", "in the Sister Study cohort",
                         "n = 50884",
                         pop = txtInt(1032),
                         .sep = "\n"),y = unit(0.5, "npc"),
  x = unit(0.2, "npc"))
included <- boxGrob(glue("Study population",
                         "n = 47810",
                         incl = txtInt(122),
                         .sep = "\n"),y = unit(0.5, "npc"),
  x = unit(0.2, "npc"))

excluded <- boxGrob(glue("Excluded:",
                         " - Withdrawal from the Sister study: n = 5",
                         " - Self-reported chemotherapy or radiotherapy for cancer prior",
                         " to baseline: n = 55",
                         " - Self-reported total thyroidectomy prior to baseline: n = 103",
                         " - Any history of invassive cancer prior to baseline: n = 2911",
                         tot = 30,
                         uninterested = 12,
                         contra = 30 - 12,
                         .sep = "\n"),
                    just = "left")

grid.newpage()
vert <- spreadVertical(eligible = eligible,
                       included = included)

excluded <- moveBox(excluded,
                    x = 0.65,
                    y =coords(vert$included)$top + distance(vert$eligible, vert$included, half = TRUE, center = FALSE))

for (i in 1:(length(vert) - 1)) {
  connectGrob(vert[[i]], vert[[i + 1]], type = "vert") %>%
    print
}

connectGrob(vert$eligible, excluded, type = "L")

# Print boxes
print(vert)
print(excluded)
rm(eligible,excluded,included,vert)

```

### Table 1: Characteristics of the study population (n=47810)

```{r}
#| label: tbl-descript
#| tbl-cap: "Characteristics of the study population (n=47810)"
#| cache: true
#| dependson: c("data", "var-list-childhood", "functions")
#| cache.extra: !expr file.mtime("C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/data/ss_earlylife_popu.csv")

table1_descript_var_label = list(
  "ident_ageexact_bl"~"Age at baseline",
  "ses_race"~"Race",
  "an_bmi_cat"~"BMI",
  "smok_status"~"Smoking status",
  "illness_thyroid_all"~"Benign thyroid disease",
  "ses_edu_level"~"Educational level",
  "ses_income"~"Household annual income",
  "ses_ADI_baseline_cat_half"~"Area depreviation index"
)
table1_descript_var_name = substr(table1_descript_var_label, 2, unlist(gregexpr("~", table1_descript_var_label)) - 3)

descript_covar = descript_table(var_name = table1_descript_var_name,
                                var_label = table1_descript_var_label,
                                data = earlylife_popu) 
rm(table1_descript_var_label,table1_descript_var_name)
descript_covar
```

## Table 2: Association between childhood and adolescent factors and thyroid cancer incidence in the Sister Study participants

```{r}
#| label: tbl-childhood
#| tbl-cap: "Association between childhood and adolescent factors and thyroid cancer incidence"
#| cache: true
#| dependson: c("data", "var-list-childhood", "functions")
#| cache.extra: !expr file.mtime("C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/data/ss_earlylife_popu.csv")

###To create Table Association between childhood and adolescent factors and thyroid cancer incidence



####Loop over childhood factors in cox models, both univariate and multivariate analyses
descript_childhood = descript_table(var_name = var_name_childhood, 
                                var_label = var_label_childhood,
                                data = earlylife_popu)|>sort_childhood()

univariate_childhood=cox_model(surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
                           var_name = var_name_childhood,
                           var_label = var_label_childhood,
                           covar = NULL,
                           data=earlylife_popu)|>sort_childhood()

multivariate_childhood=cox_model(surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
                             var_name = var_name_childhood,
                             var_label = var_label_childhood,
                             covar = covariates_childhood,
                             data=earlylife_popu)|>sort_childhood()

#assemble descriptive table and cox models' results
cox_childhood = tbl_merge(
  tbls = list(descript_childhood, univariate_childhood, multivariate_childhood),
  tab_spanner = c("**N (%)**",
                  "**Univariate models**",
                  "**Multivariate models**")) |> polish_childhood()|>modify_column_hide(c(n_event_3,exposure_3))

#return the table 
rm(descript_childhood, univariate_childhood, multivariate_childhood)
cox_childhood
```

## Table 3: Simultaneously adjusted association between thyroid cancer risk and childhood socioeconomic factors

```{r}
#| label: tbl-all-SES
#| tbl-cap: "Simultaneously adjusted association between thyroid cancer risk and childhood socioeconomic factors"

surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)"
mutual_label = list(
  "ses_income_child" ~ "Family income while growing up"         ,
  "ses_hunger_child" ~ "Ever not having enough to eat during childhood"          ,
  "ses_edu_level_child" ~ "Highest household education level at age 13",
  "ses_hh_compo_child" ~ "Household composition at age 13",
  "ses_urban_child" ~ "Childhood residence"
)
mutual_name = substr(mutual_label, 2, unlist(gregexpr("~", mutual_label)) - 3)


SES_mutual = reformulate(c(mutual_name, covariates_childhood), response = surv) |>
  coxph(method = "breslow", data = earlylife_popu) |>
  tbl_regression(
    exponentiate = TRUE,
    conf.int = TRUE,
    conf.level = 0.95,
    add_estimate_to_reference_rows = TRUE,
    label = mutual_label,
    include = mutual_name
  )|> sort_childhood() |>
  modify_fmt_fun(estimate ~ function(x)
    ifelse(is.na(x), NA, "—"),
    rows = n_event == "0") |>
  modify_fmt_fun(ci ~ function(x)
    ifelse(is.na(x), NA, "—"),
    rows = n_event == "0") |>
  modify_column_hide(p.value) |>
  modify_table_styling(
    columns = c(estimate, ci),
    rows = p.value < 0.05,
    text_format = "bold"
  ) |>
  modify_header(n_event ~ "**Number of events**",
                exposure ~ "**Person-years**",
                estimate ~ "**HR**") |>
  modify_table_body( ~ .x |> dplyr::relocate(exposure, .before = estimate))|>
  modify_footnote(abbreviation = TRUE,
                         `estimate` ~ "Multivariate models were adjusted for self-identified race/ethnicity, baseline BMI, smoking status, 
                         personal history of benign thyroid disease, educational level, household annual income, and Area Deprivation Index"
  ) 
SES_mutual
```

# Supplementary materials

## Supplementary table 1: ICD codes used to classify well-differentiated thyroid cancer

| Histological categories   | Codes                                  |
|---------------------------|----------------------------------------|
| Papillary thyroid cancer  | 8050, 8260, 8340-8344, 8350, 8450-8460 |
| Follicular thyroid cancer | 8290, 8330-8335                        |
| Unknown histology         | 8000, 8010, missing data               |

## Supplementary table 2: Association between childhood and adolescent factors and thyroid cancer incidence in the Sister Study participants, when considering only medically confirmed cases, papillary histology, competing risks, or when stratified by tumor size, cancer stage, and birth cohort

```{r}
#| label: childhood-medically-confirmed
#| output: false
#| cache: true
#| dependson: c("data", "var-list-childhood", "functions")
#| cache.extra: !expr file.mtime("C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/data/ss_earlylife_popu.csv")

cox_medical_confirmed = cox_model(
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC_source)",
  var_name = var_name_childhood,
  var_label = var_label_childhood,
  covar = covariates_childhood,
  data = earlylife_popu
)|>sort_childhood()

```

```{r}
#| label: childhood-PTC
#| output: false
#| cache: true
#| dependson: c("data", "var-list-childhood", "functions")
#| cache.extra: !expr file.mtime("C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/data/ss_earlylife_popu.csv")

cox_PTC_childhood = cox_model(
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC_PTC)",
  var_name = var_name_childhood,
  var_label = var_label_childhood,
  covar = covariates_childhood,
  data = earlylife_popu
)|>sort_childhood()

```

```{r}
#| label: childhood-cmprsk
#| output: false
##| cache: true
##| dependson: c("data", "var-list-childhood", "functions")
# ##| cache.extra: !expr file.mtime("C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/data/ss_earlylife_popu.csv")
#  earlylife_popu$ident_compete_factor = factor(earlylife_popu$ident_compete)
#  
#  cox_cmprsk_childhood = competing_risk_model(
#    surv = "Surv(ident_ageexact_bl, ident_EOF, ident_compete_factor)",
#    var_name = var_name_childhood,
#    var_label = var_label_childhood,
#    covar = covariates_childhood,
#    data = earlylife_popu
#  )|>sort_childhood()

```

```{r}
#| label: childhood-stageI
#| output: false
#| cache: true
#| dependson: c("data", "var-list-childhood", "functions")
#| cache.extra: !expr file.mtime("C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/data/ss_earlylife_popu.csv")

cox_stage_I_childhood = cox_model(
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC_stageI)",
  var_name = var_name_childhood,
  var_label = var_label_childhood,
  covar = covariates_childhood,
  data = earlylife_popu
)|>sort_childhood()

```

```{r}
#| label: childhood-advanced-stage
#| output: false
#| cache: true
#| dependson: c("data", "var-list-childhood", "functions")
#| cache.extra: !expr file.mtime("C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/data/ss_earlylife_popu.csv")

cox_adv_stage_childhood = cox_model(
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC_adv_stage)",
  var_name = var_name_childhood,
  var_label = var_label_childhood,
  covar = covariates_childhood,
  data = earlylife_popu
)|>sort_childhood()

```

```{r}
#| label: childhood-1cm
#| output: false
#| cache: true
#| dependson: c("data", "var-list-childhood", "functions")
#| cache.extra: !expr file.mtime("C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/data/ss_earlylife_popu.csv")

cox_1cm_childhood = cox_model(
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC_1cm)",
  var_name = var_name_childhood,
  var_label = var_label_childhood,
  covar = covariates_childhood,
  data = earlylife_popu
)|>sort_childhood()

```

```{r}
#| label: childhood-larger-tumor
#| output: false
#| cache: true
#| dependson: c("data", "var-list-childhood", "functions")
#| cache.extra: !expr file.mtime("C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/data/ss_earlylife_popu.csv")

cox_large_tumor_childhood = cox_model(
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC_large_tumor)",
  var_name = var_name_childhood,
  var_label = var_label_childhood,
  covar = covariates_childhood,
  data = earlylife_popu
)|>sort_childhood()

```

```{r}
#| label: childhood-strat-by-birth-cohort
#| cache: true
#| dependson: c("data", "var-list-childhood", "functions")
#| cache.extra: !expr file.mtime("C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/data/ss_earlylife_popu.csv")

cox_birth_yr_childhood = stratification_wide(
  strat_var = ident_birth_year,
  surv = "Surv(ident_ageexact_bl, ident_EOF, ident_DTC)",
  var_name = var_name_childhood,
  var_label = var_label_childhood,
  covar = covariates_childhood,
  data = earlylife_popu
) |>sort_childhood()

cox_birth_yr_childhood$table_styling$header$spanning_header =
  ifelse(
    cox_birth_yr_childhood$table_styling$header$spanning_header == "**Before 1951**",
    "**Birth cohort: Before 1951**",
    ifelse(
      cox_birth_yr_childhood$table_styling$header$spanning_header == "**1951 or after**",
      "**Birth cohort: 1951 or after**",
      cox_birth_yr_childhood$table_styling$header$spanning_header
    )
  )

```

```{r}
#| label: tbl-childhood-censor-all
#| tbl-cap: "Association between childhood and adolescent factors and thyroid cancer incidence in the Sister Study participants, when considering only medically confirmed cases, papillary histology, competing risks, or when stratified by tumor size, and cancer stage"
##| cache: true
##| dependson: c("childhood-medically-confirmed","childhood-PTC","childhood-cmprsk","childhood-stageI","childhood-advanced-stage","childhood-1cm","childhood-larger-tumor")
##| cache.extra: !expr file.mtime("C:/Users/trant4/OneDrive - National Institutes of Health/Trinh/Postdoc NCI - Risk factors and thyroid cancer/SS/data/ss_earlylife_popu.csv")

sensi_censor_childhood = tbl_merge(
  tbls = list(#cox_cmprsk_childhood, 
              cox_medical_confirmed, cox_PTC_childhood,  
              cox_stage_I_childhood, cox_adv_stage_childhood, cox_1cm_childhood, cox_large_tumor_childhood,cox_birth_yr_childhood),
  tab_spanner = c(
   # "**Competing risks**",
    "**Medically confirmed DTC cases**",
    "**Papillary thyroid cancer**",
    "**Stage I thyroid cancer**",
    "**Advanced stage thyroid cancer**",
    "**Tumor size of less than 1 cm**",
    "**Tumor size of 1 cm or more**",
    "**Birth cohort: Before 1951**"
  )
) 
#   modify_footnote(
#                   abbreviation = TRUE,
#                   `estimate_2` ~ "Multivariate models were adjusted for self-identified race/ethnicity, baseline BMI, smoking status, 
#                          personal history of benign thyroid disease, educational level, household annual income, and Area Deprivation Index"
#   ) 
#  modify_column_hide(c(exposure_1, exposure_2, exposure_3, exposure_4, exposure_5,exposure_6, exposure_7,exposure_1_8, exposure_2_8))

#sensi_censor_childhood$table_styling$header$spanning_header[213:238]=rep("**Birth cohort: 1951 or after**",26)
sensi_censor_childhood
```
